### 2019-09-06 QD_presentation.tex

\documentclass[ignorenonframetext,]{beamer}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\newif\ifbibliography
\hypersetup{
            pdftitle={The evolution of badges of status with learners},
            pdfauthor={Andrés Quiñones},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls

% Prevent slide breaks in the middle of a paragraph:
\widowpenalties 1 10000
\raggedbottom

\AtBeginPart{
  \let\insertpartnumber\relax
  \let\partname\relax
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \let\insertsectionnumber\relax
    \let\sectionname\relax
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \let\insertsubsectionnumber\relax
  \let\subsectionname\relax
  \frame{\subsectionpage}
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
\usepackage{float} \usepackage{multicol} \usepackage{amsmath}
\usepackage{graphicx} \usepackage{array}

\title{The evolution of badges of status with learners}
\author{Andrés Quiñones}
\date{}

\begin{document}
\frame{\titlepage}

\begin{frame}{The Hawk-Dove game}

Individuals have one of two genetically determined phenotypic
strategies. \emph{Hawks} are willing to start a conflict over resources,
while \emph{doves} prefer to stand down in the hope to share the
resource without an aggressive contest.

\begin{center}\includegraphics[width=0.5\linewidth]{Images//HDgame} \end{center}

\end{frame}

\begin{frame}{The hawk-dove game}

\begin{align*}
w_H &= p_H \frac{V-C}{2}+(1-p_h) V\\
w_D &= p_H 0 + (1-p_H)\frac{V}{2}
\end{align*}

\vspace{-0.8cm}

\begin{center}\includegraphics[width=0.55\linewidth]{Images//Hawk_dove_plot} \end{center}

\end{frame}

\begin{frame}{The hawk-dove game}

\begin{figure}
\includegraphics[width=0.8\linewidth]{Simulations/mutType_/basicHawkDove} \caption{\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.}\label{fig:unnamed-chunk-3}
\end{figure}

\end{frame}

\begin{frame}{What about signals?}

\begin{center}\includegraphics[height=0.8\textheight]{Images//hosp_alexandra_mackenzie} \end{center}

\end{frame}

\begin{frame}{What about signals?}

\begin{block}{When are signals honest?}

\begin{itemize}
\tightlist
\item
  Impossible to fake
\item
  Individuals have common interests
\item
  Handicap principle (signal´s cost is proportional to quality)
\item
  Social costs?
\end{itemize}

\begin{center}\includegraphics[width=1\linewidth]{Images//social_costs} \end{center}

\end{block}

\end{frame}

\begin{frame}{What about signals?}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//comSystemBot} \end{center}

\end{frame}

\begin{frame}{What about learning?}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//pavlov} \end{center}

\end{frame}

\section{Associative learning}\label{associative-learning}

\begin{frame}{Reinforcement learning theory}

\begin{columns}[T]
 \column{0.5\linewidth}
 \begin{equation*}
  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
  \end{equation*}
  \pause
  \includegraphics[width=1\textwidth]{Images//learn_curve.png}
 \column{0.5\linewidth}
  \pause
  \includegraphics[height=.9\textheight]{Images//predErrorNeur.png}
\end{columns}

\end{frame}

\begin{frame}{Reinforcement learning theory}

\begin{center}\includegraphics[width=0.9\linewidth]{Images//learnUpdate} \end{center}

\end{frame}

\begin{frame}{Environmental states}

\begin{block}{Discrete states}

\includegraphics[width=0.8\textwidth]{Images//discreteStates.png}

\pause

\end{block}

\begin{block}{Continuos states}

\includegraphics[width=0.8\textwidth]{Images//contStates.png}

\end{block}

\end{frame}

\begin{frame}{Continuos environmental states}

\begin{equation*}
  \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error}
\end{equation*}

\begin{figure}
\includegraphics[width=0.6\linewidth]{cartoonRBF} \caption{\label{fig:learning_cartoonRBF}Function approximation for the actor and the critic. Using random values for the responses of each center. Circles show the location of the centers and the maximum response they trigger, while the lines show the total response along the badge size continium.}\label{fig:fig2}
\end{figure}

\end{frame}

\section{Our model}\label{our-model}

\begin{frame}{The hawk-dove game 2.0}

\begin{center}
\begin{tabular}{ >{\centering\arraybackslash}p{1cm} | >{\centering\arraybackslash}p{4.5cm} | >{\centering\arraybackslash}p{4.5cm} }
& H & D \\ \hline
H & $p_w V\frac{-C}{2} + (1-p_w) \frac{-C}{2}$ & $V$ \\ \hline
V & $0$ & $\frac{V}{2}$\\
\end{tabular}
\end{center}

\begin{equation*}
p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
\end{equation*}

\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-9-1} \end{center}

\end{frame}

\begin{frame}{Sender Code}

\begin{equation*}
s_i = \frac{1}{1+e^{-(\epsilon_i+\gamma_iQ_i)}}
\label{eq:react_norm}
\end{equation*}

\begin{center}\includegraphics[width=0.5\linewidth]{QD_presentation_files/figure-beamer/unnamed-chunk-10-1} \end{center}

\end{frame}

\begin{frame}{Receiver code}

\begin{equation*}
  \Delta V_{t(s)}=\alpha (R_t-V_t)
\end{equation*}

\begin{center}\includegraphics[width=0.6\linewidth]{cartoonRBF} \end{center}

\end{frame}

\section{Results}\label{results}

\begin{frame}{When individuals do \emph{NOT} vary in quality}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn} \end{center}

\end{frame}

\begin{frame}{When individuals \emph{DO} vary in quality}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//hawkDoveLearn_1} \end{center}

\end{frame}

\begin{frame}{Overal effect of quality variation}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//QualStDv_//effectQualVariance} \end{center}

\end{frame}

\begin{frame}{How does variation among \emph{learners} look like?}

\begin{center}\includegraphics[width=0.7\linewidth]{Simulations//QualStDv_//WeightsVarQualSt} \end{center}

\end{frame}

\begin{frame}{How do the learning dynamics look like?}

\includegraphics[width=0.8\linewidth]{Simulations//mutType_//learnDyn200}

\end{frame}

\begin{frame}{How do learners behave when signals are honest?}

\includegraphics[width=1\linewidth]{Simulations//learHonest_//QualStDv_//weightsVarQualSt}

\end{frame}

\begin{frame}{How do learners fare against pure strategies?}

\begin{center}\includegraphics[width=0.8\linewidth]{Simulations//mutType_//evolDynTypes} \end{center}

\end{frame}

\begin{frame}{What's next?}

\begin{itemize}
\tightlist
\item
  Let reaction norm evolve, under different initial conditions.
\item
  Let learning parmeters evolve
\item
  Let the communication system co-evolve
\end{itemize}

\end{frame}

\end{document}