### 2020-06-24 manuscript_1.0.Rmd

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
  - name: Andrés E. Quiñones
    affiliation: "1"
  - name: C. Daniel Cadena
    affiliation: "1"
  - name: Olof Leimar
    affiliation: "2"
  - name: Redouan Bshary
    affiliation: "3"
address:
  - code: "1"
    address: Laboratorio de Biología Evolutiva de Vertebrados, Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
  - code: "2"
    address: Department of Zoology, Stockholm University, Stockholm, Sweden
  - code: "3"
    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
date: "`r format(Sys.time(), '%B %d, %Y')`"
bibliography: SocialMorphCue.bib
header-includes: 
  - \usepackage{float}
  - \usepackage{setspace}\doublespacing
  - \linespread{2}
  - \usepackage{lineno}
  - \linenumbers
spacing: double
abstract: My abstract which has to be long enough to take multiple
  lines otherwise one does not see the effect of single-spacing.
keywords: "My keywords"
fig_caption: yes
keep_tex: yes
fontsize: 12pt
linespread: 2
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

For social animals the outcome of interactions with conspecifics is an important 
determinant of fitness. Irrespective of whether such interactions are of cooperative or competitive nature, individual's actions will partly determine their success in the reproductive arena. However, the best action for an invidual might vary depending on it's 
own conditions, and on who they are interacting with. Thus, an individual making their action dependent on information about interacting partners is typically adaptive [@quinones_Negotiation_2016].The source of this information however is far from trivial. In some cases interacting
partners (signallers) might be 'willing' to provide accurate information, but in 
others it might be in their own interest to conceal information [@johnstone_Recognition_1997],
or to provide erroneous information. Typically, in a given context, some proportion of 
individuals will be interested in broadcasting accurate information, and some others 
in hiding it. Take for instance an interaction between two individuals where one can help the other. Given some conditions, the donor will be interested in helping a related individual. So for a relative of the donor, broadcasting kinship is advantageous; while for an unrelated
individual, conceling the lack of kinship will be better. This same pattern applies to 
many aspects social life such as finding mates, feeding offspring, dominance relationships and aggressive interactions. In dominance and aggressive interactions an important 
information to make actions dependent on is interacting partners fighting ability, 
which is sometimes refered to as Resource Holding Potential (RHP) [@parker_Assesment_1974] , or simply quality. Stable communication systems can evolve where a signal conveys quality can be 
evolutionarily stable whenever the signal imposes a fitness cost which increases with the quality of an individual

## References