### 2022-07-25 Presentations/ISBE2022.Rmd
Conference presentation done ISBE2022

---
title: | 
  | Learning can mediate the evolution of  (cheap and costly) signals of quality
author: |
  | Andrés Quiñones
institute: |
  | Universitè de Neuchâtel
bibliography: ../SocialMorphCue.bib
date: | 
  | 
  | 
  | ISBE - Stockholm 2022
output: 
  beamer_presentation: 
    fig_caption: yes
    keep_tex: yes
classoption: "aspectratio=169"
header-includes: 
  \usepackage{float}
  \usepackage{multicol}
  \usepackage{amsmath} 
  \usepackage{graphicx}
  \usepackage{array}
  \usepackage{xcolor}
  \usepackage{makecell}
---

## Acknowledgements

 ![Red1](../Images/redouan.jpg){width=40%} |  ![Daniel](../Images/danielcadena.jpg){width=55%} |
 :---: | :---:|
 *R. Bshary* |*D. Cadena*|
 ![unine](../Images/UNINE_cmjn.jpg){width=30%} | ![uniandes](../Images/logo_largo.png){width=30%} |

## Social interactions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("../../R_files/posPlots.R")
require("here")
source("../AccFunc.R")
source("../aesth.R")

```

### The need to match partners with actions

```{r,echo=FALSE,fig.align="center",out.width='40%',fig.show='hold'}

knitr::include_graphics("../Images/sexSelec.jpg")
knitr::include_graphics("../Images/dominance2.jpg")
```

## The effect of communication

### Badge of status

```{r,echo=FALSE,fig.align='center',out.width='70%',fig.show='hold'}
knitr::include_graphics("../Images/houseSparrows.png")
# knitr::include_graphics("../Images/HouseSparrow_weak.jpg")
```

## What about signals?

### The handicap principle

```{r,echo=FALSE,fig.align='center',out.height='35%',fig.show='hold'}
knitr::include_graphics("../Images/Amotz_Zahavi_profile.jpg")
knitr::include_graphics("../Images/sexualSelection.jpg")
knitr::include_graphics("../Images/Grafen.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}

badQualRang<-seq(0,1,length=1000)
alpha<-0
beta<-6
costFunc<-function(beta,alpha,x) {d<-1/(1+exp(-alpha-beta*x))}
par(plt=posPlot())
betRang<-c(5)
matplot(y=sapply(betRang,FUN = costFunc,alpha=alpha,x=badQualRang),x=badQualRang,
        type="l",cex.lab=1.5, cex.axis=2,lwd=3,lty=1,las=1,
     col=c(2,3),ylim=c(0.5,1.1),xaxt="s",yaxt="s",ylab="",xlab="") #,col=i
# lines(badgeFunc(qualRange,3,-60,1.05)~qualRange,cex.lab=1.5,cex.axis=2,lwd=3,
#       col="blue")
# legend("bottomright",legend = betRang,title = expression(beta),col=1:3,
       # lty=1)
mtext(side=1,text="Quality",line=3.5,cex=3)
mtext(side=2,text="Survival",line=3.5,cex=3)

```
\vspace{-0.3cm}
[@zahavi_Mate_1975;@grafen_Biological_1990]

<!-- ## The handicap principle -->

<!-- ### Badges of status -->

<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
<!-- knitr::include_graphics("../Images//comSystemBot.png") -->
<!-- ``` -->

<!-- [@botero_Evolution_2010] -->

<!-- ## Are bibs really badges? -->

<!-- ```{r,echo=FALSE,fig.align='left',out.width='30%'} -->
<!-- knitr::include_graphics("../Images/Sanchez-Tojar_2018.png") -->
<!-- ``` -->

<!-- [@sanchez-tojar_Metaanalysis_2018a] -->

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_0.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_1.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_2.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_3.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_5.png")
```

## What about learning?

```{r,echo=FALSE,fig.align='center',out.height='40%'}
knitr::include_graphics("../Images//pavlov.jpg")
```

```{r,echo=FALSE,fig.align='center',out.height='35%'}
knitr::include_graphics("../Images/learn_cartoon_6.png")
```

<!-- ## How to implement learning -->

<!-- ### Reinforcement learning theory -->

<!-- \begin{columns}[T] -->
<!--  \column{0.5\linewidth} -->
<!--  \begin{equation*} -->
<!--   \Delta V_{t(s)}=\alpha \underbrace{(R_t-V_t)}_\text{prediction error} -->
<!--   \end{equation*} -->
<!--   \pause -->
<!--   \includegraphics[width=1\textwidth]{../Images//learn_curve.png} -->
<!--  \column{0.5\linewidth} -->
<!--   \pause -->
<!--   \includegraphics[height=.9\textheight]{../Images//predErrorNeur.png} -->
<!-- \end{columns} -->

<!-- ## Reinforcement learning theory -->

<!-- ```{r,echo=FALSE,fig.align='center',out.width='90%'} -->
<!-- knitr::include_graphics("../Images//learnUpdate.png") -->
<!-- ``` -->

<!-- ## The Hawk-Dove game -->

<!-- Individuals have one of two genetically determined phenotypic strategies. _Hawks_ are willing to start a conflict over resources, while _doves_ prefer to stand down in the hope to share the resource without an aggressive contest. -->

<!-- ```{r,echo=FALSE,fig.align="center",out.width='50%'} -->
<!-- knitr::include_graphics("../Images//HDgame.png") -->
<!-- ``` -->

<!-- ## The hawk-dove game -->

<!-- ```{r, echo=FALSE, fig.cap="\\label{fig:HD_game}Hawk-dove game. Dashed line is the  game theoretical prediction for frequency of hawks.", out.width = '80%'} -->
<!-- knitr::include_graphics("../Simulations/mutType_/basicHawkDove.png") -->
<!-- ``` -->

## A model of communication in animal conflict

### Conflict: Hawk-Dove game

\begin{columns}
\column{0.5\linewidth}
\tiny
\begin{center}
\begin{tabular}{ >{\centering\arraybackslash}p{0.6cm} | >{\centering\arraybackslash}p{1.5cm} | >{\centering\arraybackslash}p{1.5cm} }
& Hawk & Dove \\ \hline
Hawk & $ \makecell{p_w V\frac{-C}{2}  +  \\ (1-p_w) \frac{-C}{2}}$ & $V$ \\ \hline
Dove & $0$ & $\frac{V}{2}$\\
\end{tabular}
\end{center}

\begin{equation*}
p_w=\frac{1}{1+e^{-\beta(Q_i-Q_j)}}
\end{equation*}

\column{0.5\linewidth}

```{r,echo=FALSE,fig.align='center',out.width='50%'}
par(plt=posPlot(),las=1)
plot(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 2)~
       seq(-1,1,length.out = 1000),type="l",xlab="",
     ylab = expression(p[w]),cex.lab=3,lwd=3,col="red",cex.axis=2,ylim=c(0,1))
 mtext("Difference in quality",1,line = 4,cex = 3,las = 0)
lines(logist(seq(-1,1,length.out = 1000),alpha = 0,beta = 10)~
       seq(-1,1,length.out = 1000),lwd=3,col="springgreen4")
legend("topleft",legend=c(2,10),col=c("red","springgreen4"),lwd=3,title = expression(beta),cex=2,bty="n")
```

\end{columns}

\pause

### Communication

\begin{columns}[T]

\column{0.45\linewidth}

The sender code

\tiny
\begin{equation*}
B_i = \frac{1}{1+e^{-(\textcolor{red}{\alpha_s} +\textcolor{red}{\beta_s} Q_i)}}
\end{equation*}

```{r,echo=FALSE,fig.align='center',out.width='50%'}
 par(plt=posPlot(),las=1)
 rangeBeta<-c(4,10,-4,-10)
 rangQual<-seq(0,1,by = 0.01)
 linesSender<-sapply(rangeBeta,function(x)logist(rangQual,alpha = x/2,beta = x))
 plot(logist(rangQual,alpha = 0,beta = 0)~rangQual,type="l",xlab="",
      ylab ="",cex.lab=2,lwd=2,col="grey",cex.axis=2,ylim=c(0,1))
 mtext("Quality",1,line = 4,cex = 3,las = 0)
 matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
 mtext("Badge size",2,line = 4,cex = 3,las = 0)
 legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

```

\column{0.45\linewidth}

The receiver code (learning)

\tiny
\begin{equation*}
  \Delta V_{t(s)}=\alpha (R_t-V_t)
\end{equation*}

```{r, echo=FALSE,fig.align='center', out.width = '50%'}

nCenters<-6
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
sigSq<-0.01
weights<-rep(0,nCenters)
#  Random weights
weights<-runif(nCenters,min=-1,max=1)#RBF(centers,0.5,0.05)
rangx<-seq(0,1,length=1000)

par(plt=posPlot(),las=1)
plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
     xlab="",ylab="",ylim=c(0,1),lwd=3,xaxt="s",cex.lab=2,cex.axis=2)
# test<-sapply(1:6, function(x){RBF()})
points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=3,pch=20)
# xRBFs<-sapply(centers, function(x){seq(x-0.2,x+0.2,by=0.001)})
# RBFS<-sapply(1:6, function(x){
#   logist(weights[x]*RBF(seq(centers[x]-0.2,centers[x]+0.2,by=0.001),xCenter = centers[x],
#                         sigSq = 0.01))
# })
# matlines(x=xRBFs,y=RBFS,col="grey",lty=2,lwd=2)

mtext("p(dove)",2,line = 4,cex = 3,las = 0)
mtext("Badge size",1,line = 4,cex = 3,las = 0)

```

\end{columns}

## Learning under honest signalling

<!-- Learning simulations (without evolution yet), the sender code is set as an honest signal. -->
<!-- Receivers develop a threshold like response according to their quality -->

```{r, echo=FALSE,fig.align='center', out.width = '90%'}
knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png")
```

# Results

## Learning mediates the evolution of badges of status (handicaps)

<!-- \small -->
<!-- One of the replicates of the previous slide, the badge as a handicap -->
```{r, echo=FALSE,fig.align='center', out.width='90%'}
# knitr::include_graphics("../Simulations/betCostEvol1_/evolDyn0_betCost5.png")
knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn18_betCost5.png")
```

<!-- ## Learning can mediate the evolution of Badges of Status -->
<!-- ### Individual replicates: Badges -->
<!-- \small -->
<!-- Panels show the changes in frequency distribution along evolution ($\alpha$, $\beta$, and resulting badge respectively) for the replicate shown in the previous slide -->

<!-- ```{r, echo=FALSE,fig.align='left', out.height='28%',fig.show='hold', fig.cap="alpha - beta -Badge"} -->
<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistAlpha0_betCost5.png") -->
<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBeta0_betCost5.png") -->
<!-- knitr::include_graphics("../Simulations/betCostEvol1_/evolDistBadge0_betCost5.png") -->
<!-- ``` -->

## Learning can mediate the evolution of Badges of Status
### Individual replicates: no badges
<!-- \small -->
<!-- One of the replicates where the badge did not evolve -->

```{r, echo=FALSE,fig.align='center', out.width='90%'}
knitr::include_graphics("../Simulations/betCostEvol2_/evolDyn19_betCost5.png")
```

<!-- ## Badges of Status: do they matter? -->
<!-- ### Behavioural interactions -->

<!-- \small -->
<!-- Colours correspond to replicates, each point is one snapshot on the second half of the simulations. -->
<!-- Higher costs seem to favour the signalling equilibrium -->

<!-- ```{r, echo=FALSE,fig.align='center', out.width='90%'} -->
<!-- knitr::include_graphics("../Simulations/betCostEvol1_/compClustRep_betCost5.png") -->
<!-- ``` -->

<!-- ## The effect of costs on Badges of Status -->
<!-- ### Badge size -->
<!-- \small -->
<!-- Same as previous slide but for the resulting signal. Higher costs yield smaller averga sizes and smaller ranges.  -->

<!-- ```{r, echo=FALSE,fig.align='center', out.width='80%'} -->
<!-- knitr::include_graphics("../Simulations/betCostEvol1_/CueALL.png") -->
<!-- ``` -->

<!-- ## The evolution of cheap Badges of Status -->

<!-- <!-- ### Variation among replicates --> 
<!-- <!-- \small --> 
<!-- <!-- Here simulations where the signal has no cost -->

<!-- ```{r, echo=FALSE,fig.align='center', out.width='80%'} -->
<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/evolDynALL_nIntGroup2000.png") -->
<!-- ``` -->

## The evolution of cheap badges of Status

<!-- \small -->
<!-- Here simulations where the signal has no cost. -->

<!-- There seems to be some consistency in the three type of sender codes -->

```{r, echo=FALSE,fig.align='center', out.width='90%'}
knitr::include_graphics("../Simulations/nIntGroupNormQual_/evolDyn7_nIntGroup2000.png")
```

<!-- ## The evolution of cheap Badges of Status -->
<!-- ### How often does it happen? -->
<!-- \tiny -->
<!-- Values of $\alpha$ and $\beta$ for each individual in the different replicates. Each dot an individual, -->
<!-- each panel a replicate. Individuals are sampled from populations from the second half of the simulations. -->

<!-- Individuals in some simulations split up in three groups, others in two groups;  -->
<!-- in one simulation the population does not split up -->

<!-- ```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'} -->
<!-- knitr::include_graphics("../Simulations/nIntGroupNormQual_/corrAlphBet_nIntGroup2000.png") -->
<!-- ``` -->

<!-- ## The evolution of cheap Badges of Status -->
<!-- ### Does it matter? -->
<!-- \small -->
<!-- Behavioural interactions splitting up the replicates according to the number of "types" that  -->
<!-- evolve -->

<!-- ```{r, echo=FALSE,fig.align='left', out.width='90%',fig.show='hold'} -->
<!-- knitr::include_graphics("../Simulations/nIntGroupEvol1_/BehavIntAllClusters_nIntGroup2000.png") -->
<!-- ``` -->

## Why would you want to be recognized?

\vspace{-1cm}

### The peaceful, the aggressive and the clever

- Learning initial conditions

<!-- \tiny -->
<!-- Here, I vary the aggressive tendency for unlearned individuals in these three set ups. -->

<!-- Agressive, peaceful, clever (starts at the ESS of the classic hawk-dove game) -->

```{r,out.width='45%',fig.show='hold'}
knitr::include_graphics("../Simulations/alphaAct_/alphaAct0.4learnDyn.png",)
# knitr::include_graphics("../Images/cartoonRBF_InitAct.png")
nCenters<-6
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
sigSq<-0.01
weights<-data.table(neutral=rep(0,nCenters),agressive=rep(-2,nCenters),
        peaceful=rep(2,nCenters),clever=rep(-0.69,nCenters))
rangx<-seq(0,1,length=1000)

par(plt=posPlot(numploty = 2,idploty = 2),las=1)
actor<-sapply(weights, Actor,centers=centers)

par(plt=posPlot()*0.92)
matplot(x=rangx,y=as.matrix(actor),type='l',col=c("black","#E31A1C","dodgerblue2","green4"),lty=1,
     xlab="",ylab="",ylim=c(0,1.1),lwd=5,xaxt="s",cex.lab=1.5,cex.axis=2)
mtext(text = "p(dove)",2,line = 4,cex = 2.5,las=0)
mtext(text = "Badge size",1,line = 3.5,cex = 2.5)
points(y=logist(weights$neutral,alpha = 0,beta=1),x=centers,cex=3)
text(y = actor[1,]+0.1,x=rep(0.5,4),col=c("black","#E31A1C","dodgerblue2","green4"),
     labels = names(weights),cex=2)
```

## Why would you want to be recognized?

### The peaceful, the aggressive and the clever

```{r,echo=FALSE,out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/indVarScatter0_.png")

```

## Why would you want to be recognized?

### The peaceful, the aggressive and the clever

```{r,out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/indVarScatter1_.png")

```

## Why would you want to be recognized?

### The peaceful, the aggressive and the clever

```{r,out.width='90%',fig.show='hold'}
knitr::include_graphics("../Simulations/initAct_/indVarScatter2_.png")

```

## Take-home messages

### Costly signals
- Learning can mediate costly signals

### Cheap signals
- Learning mediates the evolution of polymorphisms with "honest" and "dishonest"
individuals
- Cheap signals depend on the innate level of aggression (before learning) of individuals

## Questions?

<!-- ```{r, echo=FALSE,fig.align='center', out.height='50%',fig.show='hold'} -->
<!--   knitr::include_graphics("../Images/BizarroParrot.jpg") -->
<!--   knitr::include_graphics("../Images/Socialnet1.png") -->
<!-- ``` -->

\begin{columns}[T]
 \column{0.7\linewidth}
  \includegraphics[height=0.8\textwidth]{../Images/BizarroParrot.jpg}
 \column{0.2\linewidth}
  \includegraphics[height=1.5\textheight]{../Images//Socialnet.png}
\end{columns}

## References