### 2023-03-10 manuscript/manuscript_1.0.Rmd
New version of the discussion finished. Started addressing Daniel´s comments

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes: "```{=latex}\n  \\usepackage{float}\n  \\usepackage{unicode-math}\n
  \ \\usepackage{setspace}\\doublespacing\n  \\usepackage[backref=page]{hyperref}\n
  \ \\linespread{2}\n  \\usepackage{lineno}\n  \\linenumbers\n  \\floatplacement{figure}{H}\n
  \ \\newcommand{\\beginsupplement}{ \\setcounter{table}{0}         \\renewcommand{\\thetable}{S\\arabic{table}}\\setcounter{figure}{0}
  \\renewcommand{\\thefigure}{S\\arabic{figure}}}\n```\n"
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: "Adaptive behavioural responses often depend on qualities of the interacting
  partner of an individual. For example, when competing for resources, an individual
  might be better off escalating fights with individuals of lower quality, while restraining
  from fighting individuals of higher quality. Communication systems involving signals
  of quality allow individuals to reduce uncertainty regarding the fighting ability
  of their partners and make more adaptive behavioral decisions. However, dishonest
  individuals can destabilize such communications systems. One open question is whether
  cognitive mechanisms, such as learning, can maintain the honesty of signals, thus,
  favoring their evolutionary stability. Here we present evolutionary simulations,
  where individuals can produce a signal proportional to their quality; and learn
  along their lifetime the best response to the signal emitted by their peers. In
  these simulations, learning on the receiver side can mediate the evolution of signals
  of quality on the sender side. When the cost of the signal is proportional to the
  quality of the sender populations are only composed of honest signalers. When the
  cost is not proportional to the quality of the signaler, the population is composed
  of both honest and dishonest signalers. We argue that learning can be a general
  cognitive mechanism playing a role in a wide range of communication systems.  \n"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache=TRUE,warnings=FALSE)
require(png)
require(here)
library(dplyr)
source(here("../R_files/posPlots.r"))
source(here("AccFunc.r"))
library("scales")
```

## Introduction

The outcome of interactions with conspecifics individuals is an important 
determinant of fitness in social animals. Irrespective of whether 
such interactions are cooperative or competitive, the actions of 
interacting individuals partly determine 
their reproductive success. However, the best action for an individual 
engaging in a social interaction might vary depending on its 
own condition, and that of individuals with 
whom it interacts. Thus, acting based on information about interacting 
partners is typically adaptive [@quinones_Negotiation_2016].
Acquiring such information, however, is far from trivial. 
In some cases, interacting partners (signallers) might be 'willing' to 
provide accurate information, but in others it might be in their 
own interest to conceal information [@johnstone_Recognition_1997], 
or to provide deceptive information [@johnstone_Badges_1993]. 
Typically, in a given context, some individuals benefit 
from broadcasting accurate information, whereas others benefit 
from concealing it. Take, for instance, an interaction between two individuals 
where one can help the other. Given some costs and benefits, 
a potential donor would be interested in helping 
related individuals. Therefore, for a relative of the donor, 
broadcasting kinship would be advantageous, whereas concealing the lack 
of kinship would be better for an unrelated individual. 
Similar scenarios may apply to a variety of aspects of social 
life such as finding mates, feeding offspring, or engaging in
dominance relationships or aggressive interactions 
[@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems such as those in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals must be arbitrary, meaning they are not 
ontogenetically correlated with quality. 
Generally signals can be evolutionarily stable whenever low quality individuals 
pay a higher fitness cost for carrying the signal. 
So, it's no longer in the interest of low quality individuals 
to fake quality by producing a large badge. Signal costs inversely proportional
to quality con come about by the production costs of the signal, 
in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
<!-- should this last part of the paragraph stay ?-->
<!-- An alternative system of communication is one where individuals  -->
<!-- gather and store information about each individual they interact with.  -->
<!-- This information should reliably allow each individual to react adaptively  -->
<!-- to the quality difference with their opponent. This type of  -->
<!-- assessment strategy is often referred to as individual  -->
<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
<!-- is restricted to  certain interaction structures, because individuals  -->
<!-- are limited by cognitive abilities in how many individuals  -->
<!-- they can recognize and track. Thus, individual recognition and  -->
<!-- badges of status have been assumed to be alternative communication  -->
<!-- systems fulfilling the same function, reducing the costs of  -->
<!-- aggression in contests over resources.  -->

<!-- Furthermore, badges of status and individual recognition are expected  -->
<!-- to evolve under different interaction structures, badges of status under  -->
<!-- global interactions and individual recognition -->
<!-- under local interactions [@dale_Signaling_2001; @sheehan_There_2016].   -->
<!-- There is no clear cut reason, however, why both systems could not  -->
<!-- work together, in fact there is some evidence -->
<!-- that they can [@chaine_Manipulating_2018]. -->

Continue here!!!!

A somewhat ignored component of communication systems, in the context 
of aggressive interactions, is the cognitive aspects of the receiver module. 
Theoretical models often assume that communication systems relying on badges of 
status to have reaction norms as their 
mechanistic underpinning [@botero_Evolution_2010], with 
individuals using their own quality and the opponent's badge to determine 
whether to aggressively engage in contest. Reaction norms 
allow individuals to respond to the available information without a big
cognitive burden. That is at least in contrast to other information processing 
mechanisms like individual recognition. With individual recognition 
individuals associate cue of their peers to their quality. These association 
must be learned throughout and individual's life, thus it has the usual 
cognitive requirements of an associative learning process. An alternative view 
of badges of status is that individuals learn 
to react to them based on their experiences [@guilford_Receiver_1991], 
which would imply that receivers must learn to associate signals with 
the fighting ability of bearers just as in systems based on individual 
recognition. In cases where badges of status vary quantitatively 
(e.g. in size or intensity), fighting ability may increase monotonically 
with attributes of the badge and would be reinforced by every interaction, 
so in principle learning would be faster than in systems involving individual 
recognition in which the association between signals and their meaning would 
vary depending on the interacting partners. In any case, learning could be 
a central cognitive mechanism in both types of communication systems, 
but the role of learning in these contexts has not been thoroughly 
explored in the empirical nor theoretical literature.

Associative learning is a key cognitive mechanism that allows 
individuals to associate rewards with environmental stimuli 
and thus behave adaptively [@staddon_Adaptive_2016]. Associative 
learning exists in all major vertebrate taxa, and in many invertebrates
as well [@heyes_Simple_2012;@macphail_Brain_1982; @staddon_Adaptive_2016; 
@behrens_Associative_2008].
Theory has shown that natural selection favours these associations 
in complex environments where conditions are hard to predict 
[@dridi_Environmental_2016]. Besides its wide taxonomic and
ecological relevance, associative learning is a flexible cognitive 
mechanism whose underpinnings show interspecific variation 
[@enquist_Power_2016; @quinones_Reinforcement_2019]. Associative learning 
is not a single mechanisms, but rather a set of cognitive structures 
and processes that can vary in their scope and complexity. Presumably, 
these structures and processes have been modified 
by natural selection and could provide new explanations for behavioural 
variation. Despite all this, associative learning is 
not often included in the narrative of evolutionary explanations of 
behavioural patterns [@fawcett_Exposing_2013; @kamil_Optimal_1983; 
@mcauliffe_Psychology_2015]. Computational models of evolution 
can play an important role to overcome the lack of integration 
between learning and evolution. Reinforcement learning theory 
encompasses a series of computational methods inspired 
on the psychological and neurological mechanisms of associative 
learning [@sutton_Reinforcement_2018]. These set of algorithms allow the 
implementation of biologically realistic problems, 
capturing the essence of learning processes 
[@frankenhuis_Enriching_2018; @quinones_Reinforcement_2019]. 
Furthermore, these algorithms can be embedded in evolutionary simulations to 
generate theoretical predictions of the effect that learning can 
have in behavioural evolution [@leimar_Learning_2019;@leimar_Effects_2022].

In here we present an evolutionary model where individuals use 
associative learning to develop a tendency to behave aggressively 
(or  peacefully), in the context of competition over resources, 
depending on the quantitative morphological trait they 
perceive in their opponent (badge). Over evolutionary time 
individuals evolve the size of their badge, and whether it depends 
on their quality. Under this simple
set up, individuals can use the badge as a signal of quality. 
We use the model to assess under what conditions 
of interaction structure we expect different communication signals to evolve.

## The model

We model the evolution of signals of quality in the context of agonistic 
interactions. For simplicity we consider a population of haploid 
individuals with non-overlapping generations. Individuals are born 
every generation with a level of quality ($Q_i$, where $i$ is subscript 
of the population vector of size $N$) given by a number between zero and one, 
which is drawn from a truncated normal distribution 
$N( 0.5,\sigma)$. An individual with quality $0$ has the lowest RHP, 
while an individual with quality $1$ has the highest. As part of 
development processes individuals produce a morphological signal (badge), 
the size of such depends on their level 
of quality according to a reaction norm given by 
$B_i=1/(1+e^{\alpha_i-\beta_i Q_i})$; where $\alpha_i$ and 
$\beta_i$ are individual specific traits that determine the shape of 
the reaction norm (Fig. \ref{fig:model-struc} A). Variation in $\alpha_i$ 
and $\beta_i$ means individuals can have either uninformative (flat) or 
informative (logistic) norms. The size of the signal is constrained 
between zero and one. We assume different values of these traits 
are given by different alleles, and are inherited from mother to offspring. 
Unless, with a small probability ($\mu$), mutation changes the allelic value 
of the offspring by an amount taken from a normal distribution $TN(0,\sigma_\mu)$.
After birth individuals go through a round of 
viability selection. Individual specific survival probability is 
given by $s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})$. Where $k_1$ and $k_2$ are parameter 
values determining the shape of the survival function. 
Importantly, if $k_2=0$ survival probability is independent of quality, 
while if $k_2>0$ lower quality individuals pay a higher price for similar-sized 
badge, fulfilling the core assumption of the handicap principle 
[@botero_Evolution_2010; @grafen_Biological_1990; @johnstone_Badges_1993].

Individuals who survive engage in a series of pairwise interactions 
where they compete for resources. 
<!-- Individuals in real populations typically  -->
<!-- do not interact at random nor with all individuals in the population  -->
<!-- [@kurvers_Evolutionary_2014]. In order to allow for non-random interactions,  -->
<!-- we use the population vector to define an interaction neighbourhood  -->
<!-- for the focal individual, the size of the neighbourhood ($g$) -->
<!-- is the number of positions around the focal's from which the partner is drawn.  -->
<!-- So, if $g=N$ interactions are global; as $g$ gets smaller  -->
<!-- interactions are more local.  -->
In an interaction each individual must 
decide whether to escalate a fight or not. 
Following the classic *hawk-dove* game [@maynard-smith_Evolution_1982], 
if the focal individual fights and its partner does not, the focal gets 
as pay-off the resource of value $V$ while its partner gets nothing; 
if both individuals restrain  from fighting they split up the resource 
in half; if both decide to fight they split up the cost of fighting and 
the winner takes over the resource. We further assume that the 
probability of wining a fight for the focal individual depends on the 
difference in quality between her and her interacting partner; specifically 
it is given by $p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})$, where $k_3$ is a 
parameter defining how strong the quality difference determines the 
winning probability; and $i$ and $j$ denote the the position 
in the population vector of the focal and its interacting 
partner respectively.

The decision of whether to escalate a fight against an interacting 
partner can be dependent on the size of the partner's badge, 
the actual dependency is determined by the focal´s experiences 
through a learning process. We implement learning using 
the actor-critic approach from reinforcement learning (RL) theory [@sutton_Reinforcement_2018;@quinones_Reinforcement_2019; @leimar_Learning_2019].
Individuals estimate the reward (pay-off) expected from  interacting 
with partners of different badge sizes (the critic in RL terminology). 
After every interaction they update the estimate of reward proportionally 
to the difference between their current estimate and the observed 
reward (prediction error $\delta$) and to the speed of learning 
($\mathrm{A}$). Furthermore, they express different probabilities of 
retreating (attacking) depending on the badge size of their opponent 
(the actor in RL). They update the probability of 
retreating (attacking)  up or down depending on whether retreating 
(attacking) leads to an increase in the reward estimation. 
So, if a focal individual decides to escalate a fight against 
an individual with small badge, and that leads to an increase in 
the reward estimation, the focal individual will increase the probability 
of escalating fights with individuals of small badges in the future. 
Given that badge size is a real number between 0 and 1, there are infinitely 
many badge sizes. Thus, the reward estimation, as well as the probability 
of retreating (escalating), must be generalized across different values. 
To implement generalization we use the linear function approximation 
method based on radial basis functions [@sutton_Reinforcement_2018]. 
Specifically, we pick $c$ feature centres, which are evenly spaced values 
along the badge size interval ([0,1]). This feature centres are the same for all individuals and stored in vector $\symbf{b}$ . Each one of these features centres 
is associated with a weight for reward estimation and tendency to play 
retreat in a given interaction. The reward estimation and the tendency 
to retreat are calculated (in every interaction) as the sum of the weights 
associated with each feature centre, weighted by the response 
triggered by the feature (Fig. \ref{fig:model-struc} B, dots 
represent the feature weights) The response of each feature centre 
diminishes as a Gaussian function  with the distance between the 
feature centre and the badge size of 
the partner (Fig. \ref{fig:model-struc} B, grey line). Formally the 
reward estimation $\hat{R}$ when the focal individual ($i$) faces 
individual $j$ is given by,
\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

where $x_z$ is the weight of feature centre $z$ on the reward estimation; 
and $\theta$ is the width of the generalization function. Similarly, 
the tendency to retreat is given by the sum of feature weights associated 
with the actor, and the probability is obtained by applying a logistic 
transformation (Fig. \ref{fig:model-struc} B, black line). 
So formally, the log-odds to retreat when facing individual $j$ is

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

where $y_z$ is the weight of feature centre $z$ on the tendency to retreat.

After all their interactions, individuals in the population 
reproduce proportionally to their total pay-off $w_i$, which is a sum of the 
baseline pay-off ($w_0$) and all the pay-offs obtained throughout their 
life. Thus, the combination of natural selection and genetic drift 
changes the distribution of values in $\alpha$ and $\beta$ that segregates 
in the population, effectively changing  the badges expressed and the 
communication system.

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A the sender code: a reaction norm determines how the badge size is determined by the quality of the individual. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B the receiver code: individuals have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Generalization is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C (applies also for D) indicates the quality of the individual"}

# Top panels -  A the sender code and  B learning for the receiver code

library(dplyr)
honestBig<-readPNG(here("Images/cart_honestBig.png"))
honestSmall<-readPNG(here("Images/cart_honestSmall.png"))
dishonest<-readPNG(here("Images/cart_dishonest.png"))
par(plt=posPlot(numplotx = 2,idplotx = 1,numploty = 1,idploty = 1,upboundx = 91),
    las=1,mfrow=c(2,1),mgp=c(2,0.7,0))
  rangPars<-data.frame(honest=c(4,2),dishonest=c(0,0))
 rangQual<-seq(0,1,by = 0.01)
 linesSender<-sapply(rangPars,function(x)logist(rangQual,alpha = x[2],beta = x[1]))
 plot(0,0,type="l",xlab="",xlim=c(0,1),
      ylab ="",cex.lab=1,lwd=2,col="grey",cex.axis=1,ylim=c(0,1.1),xaxt="n")
 axis(1)
 text(x = 0.08,y=1.05,"A",cex = 1.5)
 mtext("Quality",1,line = 2,cex = 1,las = 0)
 matlines(x=rangQual,y=linesSender,lwd=4,col=colRuns,lty=1)
 mtext("Badge size",2,line = 2.3,cex = 1,las = 0)
 # include bird cartoons
 rasterImage(honestBig,0.75,0.87,0.95,1.07)
 rasterImage(honestSmall,0.05,0.01,0.25,0.21)
 rasterImage(dishonest,0.75,0.5,0.95,0.7)
 rasterImage(dishonest,0.05,0.5,0.25,0.7)
 # legend("left",legend=rangeBeta,col=colRuns,lwd=3,title = expression(beta),cex=2,bty="n")

par(plt=posPlot(numplotx = 2,idplotx = 2,numploty = 1,idploty = 1,
                 upboundx = 91),las=1,new=TRUE)
  nCenters<-6
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
sigSq<-0.01
weights<-rep(0,nCenters)
weights<-c(-5,-2,-1,1,2,5)#runif(nCenters,min=-1,max=1)
rangx<-seq(0,1,length=1000)

plot(logist(totRBF(rangx,centers,sigSq,weights),alpha = 0,beta=1)~rangx,type='l',col=1,
     xlab="",ylab="",ylim=c(0,1.1),lwd=3,xaxt="s",cex.lab=2,cex.axis=1,yaxt='n')
axis(4,cex=1)
text(x = 0.08,y=1.05,"B",cex = 1.5)
# test<-sapply(1:6, function(x){RBF()})
points(y=logist(weights,alpha = 0,beta=1),x=centers,cex=1.5,pch=20)
respon<-logist(weights[4]*RBF(rangx,centers[4],0.01))
lines(x=rangx,y=respon,col="grey",lwd=2)
axis(side = 4,at = seq(range(respon)[1],range(respon)[2],length.out = 4),
     labels = round(seq(0,1,length.out = 4),2),line = -4,col="grey")

mtext("prob. retreat",4,line = 2.3,cex = 1,las = 0)
mtext("Badge size",1,line = 2,cex = 1,las = 0)

## Lower panels - the effect of learning

scenario1<-"alphaActDishonest"
scenario2<-"alphaActHonest"

SimsDir<-here("Simulations",paste0(scenario1,"_"))
listTest<-list.files(SimsDir,full.names = TRUE)
indList<-grep("ind",listTest,value=TRUE)
# parameter values from project folder
paramName<-list.files(here("Simulations",paste0(scenario1,"_")))
# paramName<-list.files(extSimsDir,full.names = TRUE)
paramName<-grep(".json",paramName,value=TRUE)
param<-fromJSON(here("Simulations",paste0(scenario1,"_"),paramName[2]))
# fromJSON(paramName[1])
fileId<-1 # choose the replicate

indLearn<-fread(indList[fileId]) # load file
Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
             replacement = "")
nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
             replacement = "")

gener<-0#tail(indLearn[,unique(time)],2)[2]
nCenters<-param$nCenters
sigSquar<-param$sigSq
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
# nCenters<-5
# interv<-1/nCenters
# centers<-interv*0.5+interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=100)
cexAxis<-2
colorbreaksQual<-seq(0,1,length=100)
tempPop<-indLearn[time==gener]
behavTime<- tempPop[,tail(unique(nInteract),2)[1],by=indId][,min(V1)]

rangx<-seq(0,1,by=0.01)
  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
                      function(x){x=
                        logist(totRBF(rangx,
                                      centers,sigSquar,
                                      as.double(
                                        tempPop[(nInteract==behavTime
                                                 &indId==x)
                                                # &(Quality>0.8&Quality<1)
                                                ,.SD,
                                                .SDcol=grep("WeightAct",
                                                            names(tempPop),
                                                            value = TRUE)
                                                ])),alpha = 0,beta = 1)})
  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 1,upboundx = 91))
matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="prob. retreat",
          xaxt="s",yaxt="s",lty = 1,
          col=paletteMeans(100)[
            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
          lwd=3,ylim=c(0,1.1))
text(x = 0.08,y=1.05,"B",cex = 1.5)
par(new=FALSE)
color.bar.aeqp(paletteMeans(100),min =min(colorbreaksQual),
               max = max(colorbreaksQual),nticks = 3,
               cex.tit = 0.9,title = "quality",
               numplotx = 20,numploty = 10,idplotx =9,idploty = 2)

SimsDir<-SimsDir<-here("Simulations",paste0(scenario2,"_"))
listTest<-list.files(SimsDir,full.names = TRUE)
indList<-grep("ind",listTest,value=TRUE)
# parameter values from project folder
paramName<-list.files(here("Simulations",paste0(scenario1,"_")))
# paramName<-list.files(extSimsDir,full.names = TRUE)
paramName<-grep(".json",paramName,value=TRUE)
param<-fromJSON(here("Simulations",paste0(scenario1,"_"),paramName[2]))
# fromJSON(paramName[1])
fileId<-1 # choose the replicate

indLearn<-fread(indList[fileId]) # load file
Valpar<-gsub("[[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
             replacement = "")
nampar<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[fileId],"_")[[1]],1)),
             replacement = "")

gener<-0#tail(indLearn[,unique(time)],2)[2]
nCenters<-param$nCenters
sigSquar<-param$sigSq
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
# nCenters<-5
# interv<-1/nCenters
# centers<-interv*0.5+interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=100)
cexAxis<-2
colorbreaksQual<-seq(0,1,length=100)
tempPop<-indLearn[time==gener]
behavTime<- tempPop[,tail(unique(nInteract),2)[1],by=indId][,min(V1)]

rangx<-seq(0,1,by=0.01)
  dataIndsAct<-sapply(as.list(tempPop[nInteract==behavTime,indId]),
                      function(x){x=
                        logist(totRBF(rangx,
                                      centers,sigSquar,
                                      as.double(
                                        tempPop[(nInteract==behavTime
                                                 &indId==x)
                                                # &(Quality>0.8&Quality<1)
                                                ,.SD,
                                                .SDcol=grep("WeightAct",
                                                            names(tempPop),
                                                            value = TRUE)
                                                ])),alpha = 0,beta = 1)})
  # par(plt=posPlot(numploty = 4,numplotx = 4,idploty = 2,idplotx = 1))
  par(plt=posPlot(numploty = 1,numplotx = 2,idploty = 1,idplotx = 2,upboundx = 91)
      ,new=TRUE,xpd=TRUE)
matplot(x=rangx,y=dataIndsAct,type='l',xlab="",ylab="",
          xaxt="s",yaxt="n",lty = 1,
          col=paletteMeans(100)[
            findInterval(tempPop[nInteract==1700,Quality],colorbreaksQual)],
          lwd=3,ylim=c(0,1.1))
text(x = 0.08,y=1.05,"D",cex = 1.5)
text(x = -0.1,y = -0.25,labels = "Badge size")

```

## Results

We first present the outcome of simulations where we prevent the 
evolution of the sender code (by setting the mutation rate to $0$), 
and assume that all individuals in the population display either 
an uninformative or informative badge size with respect to their quality. 
These simulations show what type of receiver strategy develops through 
a learning process. Figure \ref{fig:model-struc} (C and D) shows 
the receiver strategy developed through learning; panel C is for 
receivers that faced uninformative signals, and D informative ones. 
When learners face uninformative signals, they change their 
probability of retreating depending on their own quality. Individuals 
with higher quality (red tones) after the learning process have a 
high probability of attacking; while individuals of lower quality 
(blue tones) mostly retreat from confrontations. Thus, learning splits
the population of receivers into the two classic pure strategies of 
hawks and doves. Given that we have a assumed a monomorphic population 
with unresponsive reaction norms on the signalling side, the changes 
triggered by learning only affect a small range of badge sizes. 
In contrast, when receivers face informative reaction norms 
on the side of the signaller (panel B in fig. \ref{fig:model-struc}), 
receivers use the badge size of their interacting partners to determine 
whether to retreat or attack. The relationship is given by a 
threshold-like reaction norm, where the change from retreating to 
attacking depends on the quality of the receiver. As expected, 
the higher the quality of the receiver the larger the badge size 
that triggers a retreat.

```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$). Darker areas of the background correspond to values with high frequency, while dashed lines in both panels show the mean of the distribution. X axis in the evolutionary dynamics are given in thousands of generations. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:model-struc}." ,warning=FALSE}

scenario<-"betCostEvol3"#"alphaAct"#"nIntGroupNormQual"#
runChoi<-9 #1

# Load files -------------------------------------------------------------------
# Project folder
listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
evolList<-grep("evolLearn",listTest,value=TRUE)
indList<-grep("indLearn",listTest,value=TRUE)
paramName<-list.files(here("Simulations",paste0(scenario,"_")))
# paramName<-list.files(extSimsDir,full.names = TRUE)
paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)

param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
  # fromJSON(paramName[1])
val<-2
fileId<-val
evolList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
                    evolList,value =TRUE)
indList_runs<-grep(paste0(param$namParam,param$rangParam[2]),
                   indList,value =TRUE)
evol<-do.call(rbind,lapply(evolList_runs,fread))
pop<-do.call(rbind,lapply(indList_runs, fread))

Valpar<-param$rangParam[2]

nampar<-param$namParam

nCenters<-param$nCenters
sigSquar<-param$sigSq

# Aesthetic paremeters ---------------------------------------------------------

## Calculate clustering for all generations  -------------------

# # Choose which interaction to visualize
# lastInt<-tail(pop[,unique(nInteract)],4)[2]
# # # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
# # # to get last interaction: tail(pop[,unique(nInteract)],1)
# #
# popOneInd<-pop[nInteract==lastInt]
# vars<-c("alpha","beta","Badge")
# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
#                                 Bsamples =500,iterMax = 500)
# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
#                        by=.(idClust,time,seed)]
# clusSummary[,orderClus:=0]
# invisible(capture.output(
#   for(i in 1:dim(clusSummary)[1]){
#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
#         sort(clusSummary[time==clusSummary[i,time]&
#                            seed==clusSummary[i,seed]]$meanAlph)))
# }
# ))
# 
# 
# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
# 
# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
#                              paste0("popOneInd",nampar,Valpar,".txt")))

popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
paste0("popOneInd",nampar,Valpar,".txt")))

# Get stats from the evolutionary simulations ----------------------------------
# names(evol)

cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
        "meanAlpha", "meanBeta",
        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
        "WeightAct_0","WeightCrit_0",
        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
        "WeightAct_5","WeightCrit_5")

my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
                              upIQR = fivenum(x)[4])

evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
                .SDcols = cols,by=time]

## Plot mean and IQRs among individuals for each replicate --------------------

# get the trajectories for individual runs
traitsTrajs<-dcast(evol,time~seed,
                   value.var = c("meanAlpha","meanBeta",
                                 #"meanInitCrit",#"meanFit",
                                 # "meanInitAct","sdInitCrit","sdInitAct",
                                 "sdAlpha","sdBeta","freqHH",
                                 "freqHD","freqDD"))

finReps<-evol[time==max(time),seed]

genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]

# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
cexAxis<-0.9

nInd2print<-40

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))

adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
         pal = pal_dist,
         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
         xlim=c(0,22000))

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
       bty = "n",cex=cexAxis)
legend("topleft",legend = c("A.1"),cex=cexAxis,bty = "n",inset = -0.06)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

# grey lines to show the generations shown in the upper and lower panels

matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
         lty=1,col = "grey",lwd=2)

# Variation among individuals

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
         col=colGenesLin,lty = 2,type="l",lwd=5)

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    xaxt="s",las=1,new=TRUE)

adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
       col=colGenesLin[2],lwd=2,bty = "n")
legend("topleft",legend = c("A.2"),cex=cexAxis,bty = "n",inset = -0.06)
# grey lines to show the generations shown in the upper and lower panels
matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,beta]))),
                      lty=1,col = "grey",lwd=2)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

axis(4,cex.axis=cexAxis)

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanBeta_"),runChoi)],
         col=colGenesLin[2],lty = 2,type="l",lwd=4)

# Choose time range
gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
# Plor the actor
seqYax<-c("s",rep("n",3))
# seqYlabUp<-c("Badge",rep("",3))
seqYlabUp<-c("P(dove)",rep("",3))
seqYlabDown<-c("Badge",rep("",3))
seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
seqXlabUp<-c("","Badge","")
rangQual<-seq(0,1,length.out = 50)
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=50)
count<-0
for(genC in genstoPrint){
  count<-count+1
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
                  lowboundx = 8, title = TRUE,lowboundy = 10,
                  upboundy = 87),
      xaxt="s",las=1,new=TRUE)
  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
                             .SDcols=grep("WeightAct",
                                          names(evol),value = TRUE)])
  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
               .SDcol=c(grep("WeightAct",
                             names(evol),value = TRUE),"Quality","alpha","beta",
                        "orderClus"),
               by=indId]

  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
#          beta = meansClustmp[orderClus==x,betaMean])})
# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
#          type='l',cex.axis=cexAxis,
#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
    logist(rangx,alpha = tempPop[indId==x,alpha],
           beta = tempPop[indId==x,beta])})
  matplot(x = rangx,y = reacNormInd,
          type='l',cex.axis=cexAxis,
          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
  legend("topleft",legend =  c(paste0("B.",count)),bty = "n",inset = -0.06)
  }

rm(list=grep("temp",ls(),value = T))

```

When we allow the evolution of the sender code ($\alpha$ and $\beta$ 
change subject to natural selection and genetic drift), and the badge 
size work as a handicap (i.e. the cost of the badge is inversely 
proportional to the quality of the individual), the sender code can evolve 
to produce an honest signal (Fig. \ref{fig:handicap}). The evolution of 
the sender code does not happen immediately after the start of the 
evolutionary process. Instead, the evolutionary dynamics of the reaction 
norm parameters (Fig. \ref{fig:handicap}, A) show a set of 
steps. First, the intercept of the reaction norm ($\alpha$) evolves to 
higher values, reducing the average badge size in the populations. 
That makes sense because reaction norms segregating in the population are 
flat at this point, and consequently the receiver code does not respond to 
the badge size, bigger badges are costlier and do not trigger lower attack 
probabilities. During those first generations, the slope of the sender 
reaction norm ($\beta$)  remains around the initial value of zero. 
At around generation 4000, the slope evolves toward positive values. 
Positive values in the slope imply that larger badges correlate 
with higher quality (Fig. \ref{fig:handicap} B). Receivers learn to react to such 
correlation, increasing the probability of retreat when facing 
individual of larger badges (Fig. \ref{fig:handicap} C). Hence, natural selection 
favours larger values of $\beta$, eventually leading to an evolutionary 
equilibrium where badge size is an honest signal of quality mediated
by the learned responses of receivers (Fig. \ref{fig:handicap} B.4 and C.4). 
The evolutionary trajectory portrayed in figure 
\ref{fig:handicap} A is not the only possible trajectory. If the slope of the 
sender reaction norms ($\beta$) evolves toward negative values before badges 
become handicaps, receivers never learn to react to the size of badges. 
Thus, badges are only costly and do not provide information about the 
quality of an individual (Fig. \ref{fig:noBadge}) . Eventually, the badge 
disappears from the population. In contrast, when the badge does not 
work as a handicap but instead is cost-free, the evolutionary process 
never leads to the establishment of an honest signal. 
Instead, subject to genetic drift the 
population either evolves towards the disappearance of the badge, or to the 
maximum badge size. In either of this cases reaction norms are flat, so the 
badge does not provide any quality information.

The amount of information that agents are able to collect through learning 
throughout their life time can strongly change the outcome of 
the evolutionary dynamics. In the simulations presented in figure 
\ref{fig:handicap} and \ref{fig:noBadge}, 
individuals learned with high learning rate ($\alpha = 0.4$) and interacted 
repeatedly along their lifetime (on average 2000 interactions). When we reduce 
the number of interactions that individuals have over their lifetime to 300 
interactions on average, we see an drastic increase in the morphological and genetic
variation present in the population. Fig. \ref{fig:branching} shows 
changes in the distribution of the sender reaction norm parameters
along evolutionary time, assuming the badge is cost-free. 
In panels A1 and 2, darker areas show trait values with high frequency in the population. 
Populations start monomorphic with a value of zero on both traits, 
and mutations quickly build up a normal distribution around the starting 
value. Within the first 2000 generations the unimodal distribution in 
the intercept ($\alpha$) splits into a bimodal one. 
Later in evolutionary time one of the peaks splits further, 
so at the end of the evolutionary simulation the distribution of the intercept 
($\alpha$) in the population shows three peaks. In the case of $\beta$, the 
peaks in the distribution are not so clear-cut, but it's clear that there is an 
increase in the variance of the distribution. These changes in the 
distribution of the sender reaction norm parameters 
imply that individuals can generally be classified into three 
distinct types (\ref{fig:branching}). Two types express a flat reaction 
norm with extreme values for the badge size, meaning that their badge size 
is not informative of their quality, while the third type shows 
intermediate badge sizes which are determined 
by the quality of the individual (Fig. \ref{fig:branching} B.3-4). 
In this particular simulation, the values of the slope in the informative group
are negative, which implies that badge size correlates negatively with quality. 
Furthermore, the receiver reaction norms developed through learning, particularly those 
of individuals of intermediate quality, respond to the signal of their 
sender type by increasing the probability of retreating from a fight with 
individuals with smaller badges (Fig. \ref{fig:branching}).

```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels."}

# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))

scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
runChoi<-12 #1

# Load files -------------------------------------------------------------------
# Project folder
listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
evolList<-grep("evolLearn",listTest,value=TRUE)
indList<-grep("indLearn",listTest,value=TRUE)

val<-2
fileId<-val

paramName<-list.files(here("Simulations",paste0(scenario,"_")))
paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)

param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))

evolList_runs<-grep(paste0(param$namParam,param$rangParam),
                    evolList,value =TRUE)
indList_runs<-grep(paste0(param$namParam,param$rangParam),
                   indList,value =TRUE)

evol<-do.call(rbind,lapply(evolList_runs,fread))
pop<-do.call(rbind,lapply(indList_runs, fread))

Valpar<-param$rangParam

nampar<-param$namParam

nCenters<-param$nCenters
sigSquar<-param$sigSq

# Aesthetic paremeters ---------------------------------------------------------

## Calculate clustering for all generations  -------------------

### Choose which interaction to visualize
# lastInt<-tail(pop[,unique(nInteract)],4)[2]
# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
# # to get last interaction: tail(pop[,unique(nInteract)],1)
# 
# popOneInd<-pop[nInteract==lastInt]
# vars<-c("alpha","beta","Badge")
# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
#                                 Bsamples =500,iterMax = 500)
# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
#                        by=.(idClust,time,seed)]
# clusSummary[,orderClus:=0]
# invisible(capture.output(
#   for(i in 1:dim(clusSummary)[1]){
#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
#         sort(clusSummary[time==clusSummary[i,time]&
#                            seed==clusSummary[i,seed]]$meanAlph)))
# }
# ))
# 
# 
# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
# 
# 
# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
#                              paste0("popOneInd",nampar,Valpar,".txt")))

popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
                             paste0("popOneInd",nampar,Valpar,".txt")))

# Get stats from the evolutionary simulations ----------------------------------
# names(evol)

cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
        "meanAlpha", "meanBeta",
        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
        "WeightAct_0","WeightCrit_0",
        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
        "WeightAct_5","WeightCrit_5")

my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
                              upIQR = fivenum(x)[4])

evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
                .SDcols = cols,by=time]

## Plot mean and IQRs among individuals for each replicate --------------------

# get the trajectories for individual runs
traitsTrajs<-dcast(evol,time~seed,
                   value.var = c("meanAlpha","meanBeta",
                                 #"meanInitCrit",#"meanFit",
                                 # "meanInitAct","sdInitCrit","sdInitAct",
                                 "sdAlpha","sdBeta","freqHH",
                                 "freqHD","freqDD"))

finReps<-evol[time==max(time),seed]

genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]

# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))

adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
         pal = pal_dist,
         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
         xlim=c(0,22000))

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
       bty = "n",cex=cexAxis)
legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

# grey lines to show the generations shown in the upper and lower panels

matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
         lty=1,col = "grey",lwd=2)

# Variation among individuals

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
         col=colGenesLin,lty = 2,type="l",lwd=5)

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    xaxt="s",las=1,new=TRUE)

adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
       col=colGenesLin[2],lwd=2,bty = "n")
legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
# grey lines to show the generations shown in the upper and lower panels
matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,beta]))),
                      lty=1,col = "grey",lwd=2)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

axis(4,cex.axis=cexAxis)

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanBeta_"),runChoi)],
         col=colGenesLin[2],lty = 2,type="l",lwd=4)

# Choose time range
gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
# Plor the actor
seqYax<-c("s",rep("n",3))
# seqYlabUp<-c("Badge",rep("",3))
seqYlabUp<-c("P(dove)",rep("",3))
seqYlabDown<-c("Badge",rep("",3))
seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
seqXlabUp<-c("","Badge","")
rangQual<-seq(0,1,length.out = 50)
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=50)
count<-0
for(genC in genstoPrint){
  count<-count+1
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
                  lowboundx = 8, title = TRUE,lowboundy = 10,
                  upboundy = 87),
      xaxt="s",las=1,new=TRUE)
  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
                             .SDcols=grep("WeightAct",
                                          names(evol),value = TRUE)])
  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
               .SDcol=c(grep("WeightAct",
                             names(evol),value = TRUE),"Quality","alpha","beta",
                        "orderClus"),
               by=indId]

  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
#          beta = meansClustmp[orderClus==x,betaMean])})
# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
#          type='l',cex.axis=cexAxis,
#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
    logist(rangx,alpha = tempPop[indId==x,alpha],
           beta = tempPop[indId==x,beta])})
  matplot(x = rangx,y = reacNormInd,
          type='l',cex.axis=cexAxis,
          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
  legend("topleft",legend =  c(paste0("B.",count)),bty = "n",inset = -0.06)
  }

rm(list=grep("temp",ls(),value = T))

```
The effect of a limited number of interactions on the amount of variation in 
the evolving parameters plays a role when the signal follows the handicap
principle as well. In fig \ref{fig:honestDiv}, we show simulations where individuals 
have on average 300 interactions in their life and the cost of the signal
is proportional to the quality, following the handicap principle. 
The evolutionary process leads to a combination of reaction norm parameters where 
there is a positive correlation between the size of the badge and the quality 
of the individual. This relationship however is muddled by the fact that there 
are two clusters of values for the intercept ($\alpha$) and slope ($\beta$) 
of the reaction norm in the population. Thus, there are two types 
of reaction norms. One with a steeper slope, meaning that it expresses 
a higher badge size for a given quality. This effect of an increased 
variance in the trait distribution is not only triggered by lower 
number of interactions. Larger variances are also found When we assume 
a lower speed of learning (data not shown). This suggest
that, limits to the amount of information that individuals acquire through
learning, allows the coexistence of different communication strategies 
[@botero_Evolution_2010].

```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of diverse costly signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}

# knitr::include_graphics(here("Simulations","nIntGroupNormQual_","evolDyn7_nIntGroup2000.png"))

scenario<-"betCostEvol4"#"alphaAct"#"nIntGroupNormQual"#
runChoi<-1

# Load files -------------------------------------------------------------------
# Project folder
listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
evolList<-grep("evolLearn",listTest,value=TRUE)
indList<-grep("indLearn",listTest,value=TRUE)
paramName<-list.files(here("Simulations",paste0(scenario,"_")))
# paramName<-list.files(extSimsDir,full.names = TRUE)
paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*5_*",value = TRUE)

param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[2]))
  # fromJSON(paramName[1])

val<-2
fileId<-val

evolList_runs<-grep(paste0(param$namParam,param$rangParam),
                    evolList,value =TRUE)
indList_runs<-grep(paste0(param$namParam,param$rangParam),
                   indList,value =TRUE)

evol<-do.call(rbind,lapply(evolList_runs,fread))
pop<-do.call(rbind,lapply(indList_runs, fread))

Valpar<-param$rangParam

nampar<-param$namParam

nCenters<-param$nCenters
sigSquar<-param$sigSq

# Aesthetic paremeters ---------------------------------------------------------

## Calculate clustering for all generations  -------------------

### Choose which interaction to visualize
# lastInt<-tail(pop[,unique(nInteract)],4)[2]
# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
# # to get last interaction: tail(pop[,unique(nInteract)],1)
# 
# popOneInd<-pop[nInteract==lastInt]
# vars<-c("alpha","beta","Badge")
# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
#                                 Bsamples =500,iterMax = 500)
# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
#                        by=.(idClust,time,seed)]
# clusSummary[,orderClus:=0]
# invisible(capture.output(
#   for(i in 1:dim(clusSummary)[1]){
#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
#         sort(clusSummary[time==clusSummary[i,time]&
#                            seed==clusSummary[i,seed]]$meanAlph)))
# }
# ))
# 
# 
# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
# 
# 
# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
#                              paste0("popOneInd",nampar,Valpar,".txt")))

popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
                             paste0("popOneInd",nampar,Valpar,".txt")))

# Get stats from the evolutionary simulations ----------------------------------
# names(evol)

cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
        "meanAlpha", "meanBeta",
        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
        "WeightAct_0","WeightCrit_0",
        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
        "WeightAct_5","WeightCrit_5")

my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
                              upIQR = fivenum(x)[4])

evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
                .SDcols = cols,by=time]

## Plot mean and IQRs among individuals for each replicate --------------------

# get the trajectories for individual runs
traitsTrajs<-dcast(evol,time~seed,
                   value.var = c("meanAlpha","meanBeta",
                                 #"meanInitCrit",#"meanFit",
                                 # "meanInitAct","sdInitCrit","sdInitAct",
                                 "sdAlpha","sdBeta","freqHH",
                                 "freqHD","freqDD"))

finReps<-evol[time==max(time),seed]

genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]

# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))

adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
         pal = pal_dist,
         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
         xlim=c(0,22000))

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
       bty = "n",cex=cexAxis)
legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

# grey lines to show the generations shown in the upper and lower panels

matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
         lty=1,col = "grey",lwd=2)

# Variation among individuals

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
         col=colGenesLin,lty = 2,type="l",lwd=5)

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    xaxt="s",las=1,new=TRUE)

adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
       col=colGenesLin[2],lwd=2,bty = "n")
legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
# grey lines to show the generations shown in the upper and lower panels
matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,beta]))),
                      lty=1,col = "grey",lwd=2)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

axis(4,cex.axis=cexAxis)

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanBeta_"),runChoi)],
         col=colGenesLin[2],lty = 2,type="l",lwd=4)

# Choose time range
gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
# Plor the actor
seqYax<-c("s",rep("n",3))
# seqYlabUp<-c("Badge",rep("",3))
seqYlabUp<-c("P(dove)",rep("",3))
seqYlabDown<-c("Badge",rep("",3))
seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
seqXlabUp<-c("","Badge","")
rangQual<-seq(0,1,length.out = 50)
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=50)
count<-0
for(genC in genstoPrint){
  count<-count+1
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
                  lowboundx = 8, title = TRUE,lowboundy = 10,
                  upboundy = 87),
      xaxt="s",las=1,new=TRUE)
  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
                             .SDcols=grep("WeightAct",
                                          names(evol),value = TRUE)])
  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
               .SDcol=c(grep("WeightAct",
                             names(evol),value = TRUE),"Quality","alpha","beta",
                        "orderClus"),
               by=indId]

  randInds<-sample(tempPop$indId,nInd2print, replace = FALSE)

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
#          beta = meansClustmp[orderClus==x,betaMean])})
# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
#          type='l',cex.axis=cexAxis,
#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
    logist(rangx,alpha = tempPop[indId==x,alpha],
           beta = tempPop[indId==x,beta])})
  matplot(x = rangx,y = reacNormInd,
          type='l',cex.axis=cexAxis,
          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
  legend("topleft",legend =  c(paste0("B.",count)),bty = "n",inset = -0.06)
  }

rm(list=grep("temp",ls(),value = T))

```

The behaviour expressed by naive individuals (those who have not learned) 
imposes negative-frequency dependent selection, allowing for the build up of
genetic variation in reaction norms. In the simulations, presented so far we 
assumed that individuals start with a flat behavioural reaction norm such that 
they escalate aggressively a fight with a $0.5$ probability regardless of 
the badge size of the interacting partner. To asses if this initial 
conditions of the communication system had any role 
in the build up of genetic variation, we ran a series of simulations varying 
the initial conditions of the actor module in the learning model. Specifically,
we let naive individuals have a flat reaction norm with a 1) low (peaceful),
2) high (aggressive) probability of escalating a fight and 3) a probability 
corresponding to the ESS of the original hawk-dove game (clever). 
Fig. \ref{fig:startCond} shows the distribution of values of the 
intercept ($\alpha$) and slope ($\beta$) evolved in different 
replicates of the simulations. The left-hand side panel, 
that corresponds to peaceful naive individuals, is the only one where the  
distribution of values is split in different clusters. That is, the majority of 
the variation occurs within clusters. In contrast, when naive individuals behave
either aggressively or cleverly, the badges evolve toward minimum and maximum 
values, and so the variation is driven mainly by among replicate differences. 
We can make sense of these results by realizing that individuals change their 
naive behaviour quickly in ranges of the badge size that are common in 
the population. Thus, one way to repeatedly interact with the naive behaviour
is to have a rare badge size. When the naive behaviour is peaceful, individual
with a rare badge size have a fitness advantage. This triggers 
negative-frequency dependent selection and the evolution of different types of 
badge sizes. In situations where individuals learn fast and interact repeatedly 
the strength of frequency dependent selection diminishes. The diversification 
process presented here is fully dependent on the learning process and how this
process is implemented.

```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the clever and the aggressive. Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations. The panels show the three different initial conditions for the behaviour of individuals: peaceful (blue), clever (green) and aggressive (red), see maintext for details. Colours indicate the replicate simulation. In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour."}

# knitr::include_graphics("../Simulations/initAct_/indVarScatter_.png")

scenario<-"initAct4"

# extSimsDir<-#here("Simulations",paste0(scenario,"_"))
#   paste0("e:/BadgeSims/",scenario,"_")

# Load files -------------------------------------------------------------------

listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
listTest<-grep(".txt",listTest,value = TRUE)

# (listTest<-list.files(extSimsDir,full.names = TRUE))

sdList<-grep("evol",listTest,value=TRUE)
indList<-grep("ind",listTest,value=TRUE)

inds<-rbindlist(lapply(indList,filesScenar,scenario,full.name=TRUE),fill = TRUE)

inds[,diffActWeight:=abs(WeightAct_0-WeightAct_4)]

shortSce<-gsub("[^[:alpha:]]",gsub(".txt","",tail(strsplit(indList[1],"_")[[1]],2)[1]),
               replacement = "")

# Individual variation for different parameters ------------------------------

lastInt<-tail(inds[,unique(nInteract)],2)[1]

popOneInd<-inds[nInteract==lastInt]

# vars<-c("alpha","beta","Badge")
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 5,
#                                 Bsamples =500,iterMax = 500)

# repsScen<-c(12,14,13)
repsScen<-c(5,2,7)

rangx<-seq(0,1,length=50)

list.DataReact<-lapply(popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)],
function(x){
  count<-match(x,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
  tempPop<-popOneInd[get(shortSce)==x&(seed==repsScen[count]&time==max(time))]
  tempPop<-tempPop[,.SD[.N],
                   .SDcol=c(grep("WeightAct",
                                 names(popOneInd),value = TRUE),"Quality","alpha","beta"
                            # ,
                            # "orderClus"),
                   ),
                   by=indId]
  dataIndReact<-sapply(as.list(tempPop[,indId]),
                       function(x){x=
                         sapply(rangx, 
                                function(y)
                                  do.call(logist,
                                          as.list(
                                            c(y,as.double(
                                              tempPop[indId==x,.SD,
                                                      .SDcol=c("alpha","beta")])))))})
  return(dataIndReact)
})

names(list.DataReact)<-
  popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]

yaxtAll<-c("s","n","n")
xlabAll<-c("",expression(beta[s]),"")
ylabAll<-c(expression(alpha[s]),"","")
titleAll<-c("Peaceful","Clever","Aggressive")
colorAll<-c("dodgerblue2","green4","#E31A1C")
multDiscrPallet[repsScen+1]<-colorAll
plot.new()
for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])
  # par(plt=posPlot(numploty = 2,idploty = 2,
  #                 numplotx = length(popOneInd[,unique(get(shortSce))]),
  #                 idplotx = count)+c(0,0,-0.05,-0.05),
  #     las=1,new=TRUE,
  #     yaxt=yaxtAll[match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])],
  #     cex.axis=1.5,new=TRUE)
  # plot(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&seed %in% 12:15],
  #      alpha~beta,ylab="",
  #      xlab="", pch=20,cex.lab=3,cex.axis=2,las=1,cex=2,xaxt="n",
  #      ylim=range(popOneInd[,alpha])*c(1.05,1.05),
  #      xlim=range(popOneInd[,beta]),
  #      # col=alpha(multDiscrPallet[seed+1],0.3))
  #      col=multDiscrPallet[seed+1])
  # # points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
  # #                         seed == repsScen[count]],
  # #      alpha~beta,col=multDiscrPallet[seed+1],cex=3,pch=20)
  # lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
  #       lwd=2)
  # lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
  #       lwd=2)

  par(plt=posPlot(numploty = 1,idploty = 1,
                  numplotx = length(popOneInd[,unique(get(shortSce))]),
                  idplotx = count,upboundx = 98,lowboundx = 13)
      +c(-0.03,-0.03,-0.025,-0.025),
      las=1,new=TRUE,yaxt=yaxtAll[count],
      cex.axis=1.5,new=TRUE)
  plot(data=popOneInd[get(shortSce)==PAr&time>max(time)*0.8],alpha~beta,ylab="",
       xlab="", pch=20,cex.lab=cexAxis,cex.axis=cexAxis,las=1,cex=1,
       ylim=range(popOneInd[,alpha])*c(1.1,1.05),
       xlim=range(popOneInd[,beta]),col=multDiscrPallet[seed+1])
  lines(x=c(0,0),y=range(popOneInd[,alpha]),col="grey",
        lwd=2)
  lines(y=c(0,0),x=range(popOneInd[,beta]),col="grey",
        lwd=2)
  mtext(text = xlabAll[count],side = 1,line = 2,cex = cexAxis)
  mtext(text = ylabAll[count],side = 2,line = 2,las=1,cex=cexAxis)
  mtext(text = titleAll[count],side = 3,cex = cexAxis,line = 0.2,
        col = colorAll[count])

  points(data=popOneInd[(get(shortSce)==PAr&time>max(time)*0.8)&
                          seed == repsScen[count]],
       alpha~beta,col=multDiscrPallet[seed+1],cex=2,pch=20)

}
for(PAr in popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)]){
  count<-match(PAr,popOneInd[,sort(unique(get(shortSce)),decreasing = TRUE)])

  par(new=TRUE,plt=posPlot(numplotx = 6,numploty = 4,idplotx = count*2-1,
                           idploty = 1,upboundx = 96,lowboundx =11,
                           upboundy = 98)
      +c(0.02,0.02,0.03,0.03),xpd=T,xaxt="s",yaxt="s")

  plot(x=0,y=0,type='l',xlab="", ylab="",ylim=c(0,1),xlim=c(0,1),
       xaxt="n",yaxt="n")
  polygon(x=c(par("usr")[1],par("usr")[1],par("usr")[2],par("usr")[2]),
          y=c(par("usr")[3],par("usr")[4],par("usr")[4],par("usr")[3]),
          col = "white",border=T)
  matlines(x=rangx,y=list.DataReact[[count]],
    #        col = paletteMeans(100)[
    # findInterval(tempPop[,Quality],colorbreaksQual)],
    col=alpha(multDiscrPallet[repsScen[count]+1],1),
    type='l',xlab="",
    ylab="",ylim=c(0,1),lty=1,lwd=2,new=T)
  axis(1,at = c(0,1),tick = TRUE,cex.axis=.7,padj=-1.7)
  axis(2,at =c(0,1),tick = TRUE,cex.axis=.7,hadj=-0.2)
}

```

## Discussion

We have presented here an evolutionary model where the response of a signal in a 
communication system is mediated by learning processes. Individuals 
throughout their life learn the best way to respond to a quantitative 
trait of their interactive partners. The learning process can mediate the evolution 
of an honest signal under the handicap 
principle [@grafen_Biological_1990;@zahavi_Mate_1975]. This conclusion is not 
different from the classical genetic models, where responses to the signal are
inherent to the individual. However, unlike classical genetic models,
our model shows that learning can also mediate a signal polymorphism in the 
absence of the handicap principle. Under this polymorphic equilibrium different 
types of individuals, with honest and dishonest signals, compose the population. 
The amount of information gathered through learning, as a well as the 
initial conditions of the learning process, are crucial for the emergence
and maintenance of this variation. The emergence of the polymorphism requires
a limit in the amount of information collected by individuals, as well as
peaceful behaviour in naive individuals.

Associative learning is a powerful mechanism to learn about the world, about 
social partners and about one's abilities relative to the social context. 
By associating cues and signals with fitness relevant outcomes individuals
collect information that allows them to improve their reproductive potential. 
This is very obvious in classical examples such as when animals learn to avoid 
food that makes them sick, or use environmental cues to find food. 
More recent ideas have highlighted the potential role that associations 
can have in a social contexts such as: hierarchy formation 
[@leimar_Reproductive_2022;@leimar_Effects_2022],
and cooperation [@leimar_Learning_2019;@dridi_Learning_2018]. In this 
theoretical examples individuals use various sources of reward to make
adaptive choices in a given social context. Here we extend this logic to the 
evolution of a communication system mediated by a quality signal. 
Previous evolutionary models of badges of status assumed that individuals 
responded using a behavioural reaction norm, where the opponent's badge 
and individual's own quality determined the behaviour [@botero_Evolution_2010]. 
However, there isn't a clear mechanism to justify the assumption that 
individuals inherently know their own quality, particularly relative to their peers.
In our model, individuals not only learn about the quality signal, but also 
learn about their own quality relative to others. This can be seen throughout
our simulations in the responses developed by individuals of different quality. 
Because individuals learn about their own quality, 
even in the absence of an honest signal, they are able to make adaptive choices. 
Thus the learning process we have modelled and the response it mediates 
has fitness relevant consequences even in the absence of an honest signal.

Learning processes collecting information on a population wide level promote 
frequency dependent selection, in the communication system we model frequency 
dependency translates to the coexistence of different communication strategies. 
The frequency-dependence triggered by learning process was highlighted by the  
classic study of [@bond_Visual_2002] where live predators drive the evolution
of *in silico* polymorphic prey. The key to the evolution of this polymorphisms is 
that predators learn to discriminate better preys that are frequently encountered.
Thus, preys found in low frequency have a selective advantage. 
An analogous process occurs in our simulations, the learning algorithm of the 
receiver individual learns the best response for herself towards individuals 
that have a badge size common in the population. When an individual has a
rare badge size it can be favoured or unfavoured by the naive behaviour. When 
the naive behaviour is peaceful, polymorphism is promoted, when the naive 
behaviour is aggressive polymorphism is prevented. These two examples show how
behaviours dependent on learning processes seem to generally trigger frequency 
dependent selection. That is because the learning algorithm collects more 
information on the frequent values of the the trait distribution. For the sake  
of simplicity we modelled a single dimension (badge size), however this effect 
could potentially be more important when individuals learn from 
multidimensional signals.

Are bibs real signals of quality? if so are they handicaps or badge of status? 
Originally @rohwer_Social_1975a proposed the idea that certain morphological 
traits could be used by animals as status signals or signals of quality. 
However, from an evolutionary perspective it is unclear how the stability of 
such communication systems can be maintained. One option is that honesty is 
maintained by the cost of signal production; that is, when the signal works as a 
handicap [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993]. 
Alternatively, if the signal does not carry production costs inversely 
proportional to quality it may work as a convention.
In this later case, honesty is supposedly 
maintained by the aggressive reaction of receivers when the convention is 
broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. Our model, portraits 
both kinds of signals, although with some nuances. When we assume the handicap 
principle, we get the evolution of honest signals mediated by the learned 
response. When we assume cost-free signal, under certain conditions the learned 
response mediated the emergence of phenotypic variation. This variation, 
particularly in the mid-range of the distribution facilitates the establishment 
of a convention, in this range individuals respond appropriately to the trait of their 
peer. The house sparrow (*Passer domesticus*) has been a text-book example 
of badges of status. According to the prevailing narrative the bib size 
in males is a signal of the dominance rank and quality.
However, a recent meta-analysis called into question 
this narrative by showing that the effect size of the association between bib 
size and dominance rank is small and uncertain [@sanchez-tojar_Metaanalysis_2018a]. 
The simple straight correlation between bib size and dominance rank
( and quality) is expected under the handicap principle, as it is shown here and 
elsewhere [@botero_Evolution_2010;@grafen_Biological_1990;@johnstone_Badges_1993].
But, the diversification process presented here shows that conventions can 
emerge in a less straightforward way when mediated by learning processes. This 
nuances could potentially make sense of seemingly contradictory evidence on the 
correlation between plumage traits and quality. For example, conventions in 
our simulations arose only in an intermediate range of the signalling trait. 
While correlations are often performed across the whole range of variation.

The focus of theoretical and empirical work on communication systems, 
and particularly badges of status, is often explaining the presence and 
absence of certain morphological traits within populations. However, the 
morphological traits hypothesized to play a role in communication systems 
regularly present interesting patters at the phylogeographic
level. One example of this is the leapfrog pattern [@remsen_High_1984],
whereas a certain morphological trait alternates its presence and 
absence in a set of geographically adjunct populations. Patterns of 
molecular variation along geographic clines suggest that natural selection 
is involved in the emergence of these patterns 
[@cadena_Testing_2011;@cadena_Evolutionary_2007]. However, it is unclear what 
type of ecological process is behind these selective regimes. 
Given the wide variation in outcomes found in our model, that depend on both 
stochasticity and cognitive parameters, we think variation in cognition and 
learning in particular could provide some explanatory power in this respect. 
One first step, for example is to evaluate the way individuals in different 
populations respond to novel traits [@avendano_Territorial_2021].

We have presented here an evolutionary model where the evolution of a 
communication system is mediated through the learned responses of
receivers. This is a novel way to understand the evolution of communication
that integrates classical cognitive process in learning with 
evolutionary explanations of communication. This approach contributes to the 
further integration of proximate and ultimate explanations in behavioral and 
evolutionary biology.

# Supplementary material

```{=tex}
\beginsupplement
```

```{r noBadge,echo=FALSE,warning=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}

# knitr::include_graphics(here("Simulations","betCostEvol1_","evolDyn5_betCost5.png"))

scenario<-"betCostEvol3"#"alphaAct"#"nIntGroupNormQual"#
runChoi<-12 #1

# Load files -------------------------------------------------------------------
# Project folder
listTest<-list.files(here("Simulations",paste0(scenario,"_")),full.names = TRUE)
evolList<-grep("evolLearn",listTest,value=TRUE)
indList<-grep("indLearn",listTest,value=TRUE)

val<-2
fileId<-val

paramName<-list.files(here("Simulations",paste0(scenario,"_")))
paramName<-grep(".json",paramName,value=TRUE) %>% grep(pattern = "*s0_*",value = TRUE)

param<-fromJSON(here("Simulations",paste0(scenario,"_"),paramName[1]))

evolList_runs<-grep(paste0(param$namParam,param$rangParam),
                    evolList,value =TRUE)
indList_runs<-grep(paste0(param$namParam,param$rangParam),
                   indList,value =TRUE)

evol<-do.call(rbind,lapply(evolList_runs,fread))
pop<-do.call(rbind,lapply(indList_runs, fread))

Valpar<-param$rangParam[1]

nampar<-param$namParam

nCenters<-param$nCenters
sigSquar<-param$sigSq

# Aesthetic paremeters ---------------------------------------------------------

## Calculate clustering for all generations  -------------------

runChoi<-9#3
## Choose which interaction to visualize
# lastInt<-tail(pop[,unique(nInteract)],4)[2]
# # pop[,max(nInteract),by=.(seed,time)][,min(V1)]
# # to get last interaction: tail(pop[,unique(nInteract)],1)
# 
# popOneInd<-pop[nInteract==lastInt]
# vars<-c("alpha","beta","Badge")
# # ,"Badge",grep("WeightAct",names(popFinal),value = TRUE))
# popOneInd$idClust<-get_clusters(popOneInd,vars,k.max = 10,
#                                 Bsamples =500,iterMax = 500)
# clusSummary<-popOneInd[,.(meanAlph=mean(alpha),meanBet=mean(beta)),
#                        by=.(idClust,time,seed)]
# clusSummary[,orderClus:=0]
# invisible(capture.output(
#   for(i in 1:dim(clusSummary)[1]){
#   set(clusSummary,i,6L,match(clusSummary[i,meanAlph],
#         sort(clusSummary[time==clusSummary[i,time]&
#                            seed==clusSummary[i,seed]]$meanAlph)))
# }
# ))
# 
# 
# popOneInd<-merge(popOneInd,clusSummary,all = TRUE)
# 
# 
# fwrite(popOneInd,file = here("Simulations",paste0(scenario,"_"),
#                              paste0("popOneInd",nampar,Valpar,".txt")))

popOneInd<-fread(here("Simulations",paste0(scenario,"_"),
                             paste0("popOneInd",nampar,Valpar,".txt")))

# Get stats from the evolutionary simulations ----------------------------------
# names(evol)

cols<-c("freqGenHawks","freqGenDove",  "freqGenEval",  #"freqGenLearn",
        "freqFenHawks", "freqFenDoves", "freqHH", "freqHD", "freqDD", "meanCue",
        "meanAlpha", "meanBeta",
        #"meanFit", #"meanInitCrit", "sdInitCrit", "meanInitAct", "sdInitAct",
        "WeightAct_0","WeightCrit_0",
        "WeightAct_1","WeightCrit_1","WeightAct_2",  "WeightCrit_2",
        "WeightAct_3",  "WeightCrit_3", "WeightAct_4",  "WeightCrit_4",
        "WeightAct_5","WeightCrit_5")

my.summary<- function(x) list(mean = mean(x), lowIQR = fivenum(x)[2], 
                              upIQR = fivenum(x)[4])

evolStats<-evol[, as.list(unlist(lapply(.SD, my.summary))), 
                .SDcols = cols,by=time]

## Plot mean and IQRs among individuals for each replicate --------------------

# get the trajectories for individual runs
traitsTrajs<-dcast(evol,time~seed,
                   value.var = c("meanAlpha","meanBeta",
                                 #"meanInitCrit",#"meanFit",
                                 # "meanInitAct","sdInitCrit","sdInitAct",
                                 "sdAlpha","sdBeta","freqHH",
                                 "freqHD","freqDD"))

finReps<-evol[time==max(time),seed]

genstoPrint<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]

# Evolutionary dynamics for alpha - intercept of the sender --------------------------------------------
par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2,idplotx = 1,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    mfrow=c(1,1), xaxt="s",las=1,mgp=c(2,0.5,0))

adjsYlims<-abs(range(pop[seed==runChoi,alpha])[1]-
                 range(pop[seed==runChoi,alpha])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "alpha",nbins = 10,
         pal = pal_dist,
         nlevels=10,cexAxis = cexAxis,xlab="",ylab = "",xaxt="n",yaxt = "s",
         ylim = range(pop[seed==runChoi,alpha])+adjsYlims,
         xlim=c(0,22000))

mtext(text = "Trait value",cex = cexAxis,line = 1,side = 2,las=0)
legend("topright",legend = c(expression(alpha[s])),col=colGenesLin[1],lwd=2,
       bty = "n",cex=cexAxis)
legend("topleft",legend = c("A"),cex=cexAxis,bty = "n",inset = -0.06)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

# grey lines to show the generations shown in the upper and lower panels

matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,alpha]),4),nrow = 2),
         lty=1,col = "grey",lwd=2)

# Variation among individuals

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanAlpha_"),runChoi)],
         col=colGenesLin,lty = 2,type="l",lwd=5)

par(plt=posPlot(numploty = 3,idploty = 3,numplotx = 2, idplotx = 2,
                lowboundx = 8, title = TRUE,lowboundy = 15,
                upboundy = 99),
    xaxt="s",las=1,new=TRUE)

adjsYlims<-abs(range(pop[seed==runChoi,beta])[1]-
                 range(pop[seed==runChoi,beta])[2])*c(-0.1,0.2)

evolDist(indData = pop[seed==runChoi],variable = "beta",nbins = 10,pal = pal_dist,
         nlevels=10,cexAxis = 1.5,xlab="",ylab = "",xaxt="n",yaxt = "n",
         ylim = range(pop[seed==runChoi,beta])+adjsYlims,xlim=c(0,22000))
legend("topright",legend = c(expression(beta[s])),cex=cexAxis,
       col=colGenesLin[2],lwd=2,bty = "n")
legend("topleft",legend = c("B"),cex=cexAxis,bty = "n",inset = -0.06)
# grey lines to show the generations shown in the upper and lower panels
matlines(x=matrix(rep(evolStats[genstoPrint,time],each=2),nrow=2),
         y=matrix(rep(range(pop[seed==runChoi,beta]))),
                      lty=1,col = "grey",lwd=2)

axis(side=1,padj = -2,at = axTicks(1)[2:(length(axTicks(1)))],
     labels = axTicks(1)[2:(length(axTicks(1)))]/1000,cex.axis=cexAxis,
     hadj = 1.3,line = -0.5,tick = FALSE)

axis(4,cex.axis=cexAxis)

matlines(x=traitsTrajs[,time],
         y=traitsTrajs[,.SD,
                       .SDcol=paste0(c("meanBeta_"),runChoi)],
         col=colGenesLin[2],lty = 2,type="l",lwd=4)

# Choose time range
gen2plot<-round(seq(1,length(unique(evolStats$time)),length.out = 5))[2:5]
# Plor the actor
seqYax<-c("s",rep("n",3))
# seqYlabUp<-c("Badge",rep("",3))
seqYlabUp<-c("P(dove)",rep("",3))
seqYlabDown<-c("Badge",rep("",3))
seqXlabDown<-c("","Quality","")#paste0("seed: ",runChoi))
seqXlabUp<-c("","Badge","")
rangQual<-seq(0,1,length.out = 50)
interv<-1/(nCenters-1)
centers<-interv*seq(0,nCenters-1)
rangx<-seq(0,1,length=50)
count<-0
for(genC in genstoPrint){
  count<-count+1
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 1,
                  lowboundx = 8, title = TRUE,lowboundy = 10,
                  upboundy = 87),
      xaxt="s",las=1,new=TRUE)
  weightsAct<-as.double(evol[(time==unique(time)[genC])&seed==runChoi,.SD,
                             .SDcols=grep("WeightAct",
                                          names(evol),value = TRUE)])
  tempPop<-popOneInd[(time==unique(time)[genC]&seed==runChoi),.SD[.N],
               .SDcol=c(grep("WeightAct",
                             names(evol),value = TRUE),"Quality","alpha","beta",
                        "orderClus"),
               by=indId]

  randInds<-sample(tempPop$indId, nInd2print, replace = FALSE)

  dataIndAct<-sapply(as.list(tempPop[indId %in%randInds,indId]),
                     function(x){x=
                       logist(totRBF(rangx,
                                     centers,sigSquar,
                                     as.double(
                                       tempPop[indId==x,.SD,
                                               .SDcol=grep("WeightAct",
                                                           names(tempPop),
                                                           value = TRUE)
                                               ])),alpha=0,beta = 1)})
  matplot(x=rangx,y=dataIndAct,
          yaxt=seqYax[count],ylab="",xlab="",type="l",cex.lab=cexAxis,
          lwd=2,xaxt="n",ylim=c(-0.06,1.1),col = paletteMeans(100)[
            findInterval(tempPop[indId %in%randInds,Quality],
                         colorbreaksQual)],lty = 1,
          cex.axis=cexAxis)
  axis(side = 1,cex.axis=cexAxis,padj = 0)
  mtext(text = seqYlabUp[count],cex = cexAxis,line = 1.7,side = 2,las=0)
  mtext(text = seqXlabUp[count],cex = cexAxis,line = 1.2,side = 1 ,las=0)
  lines(logist(totRBF(rangx,centers,sigSquar,weightsAct),alpha=0,beta=1)~rangx,
        col=1,lwd=3,lty=2)
  legend("topleft",legend = c(paste0("C.",count)),bty = "n",inset = -0.06)
  # text(x=0.5,y=-0.015,labels = paste0("time=",unique(evolStats$time)[genC]/1000),
  #      cex=2)
  par(plt=posPlot(numplotx = 4,numploty = 3,idplotx = count,idploty = 2,
                  lowboundx = 8,lowboundy = 15,
                  upboundy = 99), las=1,new=TRUE)

# meansClustmp<-tempPop[,.(alphaMean=mean(alpha),betaMean=mean(beta)),by=orderClus]
# reacNormpClust<-sapply(meansClustmp[,orderClus], function(x){
#   logist(rangx,alpha = meansClustmp[orderClus==x,alphaMean],
#          beta = meansClustmp[orderClus==x,betaMean])})
# matplot(x = rangx,y = reacNormpClust,col=colRuns[meansClustmp$orderClus],
#          type='l',cex.axis=cexAxis,
#          xlab="",ylab="",ylim=c(0,1.1),lty=1,
#          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  reacNormInd<-sapply(tempPop[indId %in%randInds,indId], function(x){
    logist(rangx,alpha = tempPop[indId==x,alpha],
           beta = tempPop[indId==x,beta])})
  matplot(x = rangx,y = reacNormInd,
          type='l',cex.axis=cexAxis,
          xlab="",ylab="",ylim=c(0,1.1),lty=1,col=1,
          lwd=2,yaxt=seqYax[count],xaxt="s",cex.lab=cexAxis)
  mtext(text = seqYlabDown[count],cex = cexAxis,line = 1.5,side = 2,las=0)
  mtext(text = seqXlabDown[count],cex = cexAxis,line = 1.3,side = 1,las=0)
  legend("topleft",legend =  c(paste0("B.",count)),bty = "n",inset = -0.06)
  }

rm(list=grep("temp",ls(),value = T))

```

\newpage
\small

| Symbol |      Description |
|:-------:|:--------------------------------|
| $Q_i$ | Quality of individual $i$ |
| $N$   | Population size |
| $\sigma$ | Standard deviation of the truncated normal distribution from which quality is drawn |
| $B_i$ | Badge size of individual $i$ |
| $\alpha_i$ | Intercept of the badge size reaction norm for individual $i$ |
| $\beta_i$  | Slope of the badge size reaction norm for individual $i$ |
| $\mu$     | Mutation rate | 
| $\sigma_\mu$ |Standard deviation of the normal distribution from which quality is mutations are drawn |
| $s_i$ | Survival probability of individual $i$ |
| $k_1$ | Intercept of the survival function |
| $k_2$ | Slope of the survival function |
| $V$   | Value of the contested resource |
| $C$   | Cost of an escalated fight |
| $p_{ij}$ | Probability that individual $i$ wins an escalated fight against individual $j$ |
| $k_3$ | Parameter determining how important is quality defining the probability of wining |
| $\delta$ | Prediction error |
| $A$ | Speed of learning | 
| $c$ | Number of feature centres |
| $b_z$ | badge size for feature centre $z$  | 
| $\hat{R}$ | Reward estimate | 
| $x_z$ | weight of feature centre $z$ on the reward estimation |
| $\theta$ | width of the generalization function | 
| $q_{ij}$ | probability that individual $i$ retreats from a fight in an encounter with individual $j$ |
| $y_z$ | weight of feature centre z on the tendency to retreat |
| $w_i$ | total pay-off of individual $i$ |
| $w_0$ | base line pay-off |

Table: Notation of the model parameters and variables

## References