### 2023-09-14 manuscript/Title_page.Rmd
Title_page rmd file and readme md file, getting readu for realease and submission

---
title: |
   \large \textbf{Title:} The role of learning in the evolution of status signalling: a modeling approach
subtitle: |
  \normalsize \textbf{Abbreviated title:} Learning and evolution in status signalling
author: 
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: 
      - a
      - b
    corresponding: andresqp@gmail.com 
    fulladress: Rue Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486
  - name: Redouan Bshary
    affiliation: a
  - name: C. Daniel Cadena
    affiliation: 
      - b
address:
  - code: a
    address: Institute of Biology, University of Neuchatel, Neuchatel, Switzerland
  - code: b
    address: Laboratorio de Biología Evolutiva de Vertebrados, Departamento  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia
output:    
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    template: template.latex
---

## Funding
This work was supported by the posdoctoral assistant funding from the Depatamento de Ciencias Biológicas from the Universidad de los Andes granted to A.Q. and by the by the Swiss National Science Foundation grants (grant number: 310030_192673) to R.B.

## Acknowledgements
AQ and CDC kindly thank the members of Evolvert, in particular
Jorge Avendaño for pictures of the *Arremon* and discussion on
signal evolution. AQ thanks C. Botero for fruitful discussions, and thecluster of the Institute of Computer Science of the University of Neuchâtel (IIUN) for the computing resources provided.