### 2023-09-14 manuscript/Title_page.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={ The role of learning in the evolution of status signalling: a modeling approach},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\large \textbf{Title:} The role of learning in the evolution of
status signalling: a modeling approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\normalsize \textbf{Abbreviated title:} Learning and evolution
in status signalling}
\date{}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\newcommand{\getAff}[1]{
  \getIndex{#1}{a,b}
}

\begin{document}
\maketitle
\begin{flushleft}

Andrés E.
Quiñones\textsuperscript{\getAff{a}, \getAff{b}}\textsuperscript{*},
Redouan Bshary\textsuperscript{\getAff{a}},
C. Daniel Cadena\textsuperscript{\getAff{b}}\\
\bigskip
\textbf{\getAff{a}}Institute of Biology, University of Neuchatel,
Neuchatel, Switzerland\\
\textbf{\getAff{b}}Laboratorio de Biología Evolutiva de Vertebrados,
Departamento de Ciencias Biológicas, Universidad de los Andes, Bogotá,
Colombia\\
\bigskip
* Corresponding author: \href{mailto:andresqp@gmail.com}{\nolinkurl{andresqp@gmail.com}}, Rue
Emile-Argand 11, CH-2000, Neuchatel, Switzerland, tel. +41325586486\\
\end{flushleft}

\hypertarget{funding}{%
\subsection{Funding}\label{funding}}

This work was supported by the posdoctoral assistant funding from the
Depatamento de Ciencias Biológicas from the Universidad de los Andes
granted to A.Q. and by the by the Swiss National Science Foundation
grants (grant number: 310030\_192673) to R.B.

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

AQ and CDC kindly thank the members of Evolvert, in particular Jorge
Avendaño for pictures of the \emph{Arremon} and discussion on signal
evolution. AQ thanks C. Botero for fruitful discussions, and thecluster
of the Institute of Computer Science of the University of Neuchâtel
(IIUN) for the computing resources provided.

\end{document}