### 2023-09-14 manuscript/manuscript_1.0.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
  \AddToHook{env/abstract/before}{\vspace*{-4.5cm}}
  \usepackage{float}
  \usepackage{graphicx}
  \usepackage{unicode-math}
  \usepackage{setspace}\doublespacing
  \usepackage[backref=page]{hyperref}
  \linespread{2}
  \usepackage{lineno}
  \linenumbers
  \floatplacement{figure}{H}
  \newcommand{\beginsupplement}{ \setcounter{table}{0}            
  \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={ The role of learning in the evolution of status signalling: a modeling approach},
  pdfkeywords={My keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{\large \textbf{Title:} The role of learning in the evolution of
status signalling: a modeling approach}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{\normalsize \textbf{Abbreviated title:} Learning and evolution
in status signalling}
\date{}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\newcommand{\getAff}[1]{
  \getIndex{#1}{}
}

\begin{document}
\maketitle
\begin{flushleft}
\end{flushleft}
\begin{abstract}
Adaptive behavioural responses often depend on qualities of the
interacting partner of an individual. For example, when competing for
resources, an individual might be better off escalating fights with
individuals of lower quality, while restraining from fighting
individuals of higher quality. Communication systems involving signals
of quality allow individuals to reduce uncertainty regarding the
fighting ability of their partners and make more adaptive behavioral
decisions. However, dishonest individuals can destabilize such
communications systems. An open question is whether cognitive
mechanisms, such as learning, can maintain the honesty of signals, thus
favoring their evolutionary stability. We present evolutionary
simulations where individuals can produce a signal proportional to their
quality and learn along their lifetime the best response to the signals
emitted by their peers. Our simulations replicate previous results where
the handicap principle mediates the evolution of signals. In our
simulations learning on the receiver side can mediate the evolution of
signals of quality on the sender side. When the cost of the signal is
proportional to the quality of the sender, all individuals in
populations are honest signalers. In contrast to traditional models
which predict the absence of signals when the cost is not proportional
to the quality of the signaler, our model revealed that learning
facilitates the evolution of a polymorphism in which populations
comprise both honest and dishonest signalers. We argue that learning may
have a role in the evolution and dynamics of a wide range of
communication systems and more generally in behavioral responses.
\end{abstract}
\subsection*{Lay summary}
Biological signals allow individuals to respond appropietly. For
example, a singal of fighting hability would allow peers to decide
whether to fight another individual. But, what prevents a weaklings from
faking to be strong? Using a theoretical approach we assess whether
learning has a role in maintaining signals honest. We find that when
individuals learn how best to respond to a signal, populations are
composed of a mixed of honest and dishonest signals.

\newpage

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

The outcome of interactions with conspecific individuals is a critical
determinant of fitness in social animals. Irrespective of whether such
interactions are cooperative or competitive, the actions of interacting
individuals influence their reproductive success. However, the best
action for an individual engaging in a social interaction might vary
depending on its own condition and that of individuals with whom it
interacts. Thus, acting based on information about interacting partners
is typically adaptive (Quiñones et al. 2016), yet acquiring such
information is far from trivial. In some cases, interacting partners
(\emph{i.e.} signallers) might be `willing' to provide accurate
information, but in others it might be in their own interest to conceal
information (Johnstone 1997) or to provide deceptive information
(Johnstone and Norris 1993). Typically, in a given context, some
individuals benefit may from broadcasting accurate information, whereas
others may benefit from concealing it. Take, for instance, an
interaction between two individuals where one can help the other. Given
some costs and benefits, a potential donor is likely to be interested in
helping related individuals. Therefore, for a relative of the donor,
broadcasting kinship would be advantageous, whereas concealing the lack
of kinship would be better for an unrelated individual. Similar
scenarios may apply to a variety of aspects of social life such as
finding mates, feeding offspring, or engaging in dominance relationships
or aggressive interactions (Bradbury and Vehrencamp 2011; Møller 1988;
Tibbetts and Dale 2004).

In dominance and aggressive interactions, a crucial piece of information
to guide individual actions is the fighting ability of interacting
partners, which is sometimes referred to as Resource Holding Potential
(RHP, Parker 1974) or simply as quality. Responsiveness to the quality
of an opponent is central to communication systems in which individuals
exhibit badges of status, where a signal typically conveys quality
(Johnstone and Norris 1993; Rohwer 1975). Such signals tend to be
arbitrary, however, meaning the signals could potentially be produced by
high- and low-quality individuals alike, opening up the possibility of
dishonest signalling and raising the question of evolutionary stability.
Signals can be evolutionarily stable whenever low-quality individuals
pay a higher fitness cost for carrying the signal such that it is not in
the interest of low-quality individuals to fake quality by producing a
large badge. Signal costs inversely proportional to quality may come
about as a result of the production costs of the signal, in which case
the signal works like a handicap (Botero et al. 2010; Johnstone and
Norris 1993; Grafen 1990; Zahavi 1975). Alternatively, the costs can
come from the aggressive reaction of receivers when the convention of
the communication is broken (Enquist, Ghirlanda, and Hurd 2010; Tibbetts
and Dale 2004). For example in paper wasps, subordinate individuals with
experimentally manipulated dominant-like facial patterns received more
aggression from the dominant (Tibbetts and Dale 2004). In this case,
costs are socially rather than developmentally mediated and are
triggered by the violation of the convention, although how the
convention is established in the first place is unclear.

A somewhat ignored component of communication systems in the context of
aggressive interactions are cognitive aspects of the receiver module.
Theoretical models often assume that communication systems relying on
badges of status have reaction norms as their mechanistic underpinning
(Botero et al. 2010), with individuals using their own quality and the
badges of opponents to determine whether to aggressively engage in
contests. Reaction norms allow individuals to respond to the available
information without a large cognitive burden because the computations
involved in the reaction norm response do not require much memory nor
advanced algorithmic processes. This contrasts with other
information-processing mechanisms like individual recognition where
individuals associate cues of their peers to their quality. Because
these associations must be learnt throughout the life of individuals,
they have the usual cognitive requirements of associative learning
processes. An alternative view of badges of status is that individuals
learn to react to them based on their experiences (Guilford and Dawkins
1991), which would imply that receivers must learn to associate signals
with the fighting ability of bearers just as in systems based on
individual recognition. In cases where badges of status vary
quantitatively (\emph{e.g.} in size or intensity), fighting ability may
increase monotonically with attributes of the badge and would be
reinforced by every interaction. Therefore, in principle, learning in
such scenarios would be faster than in systems involving individual
recognition in which the association between signals and their meaning
varies depending on the interacting partners. In any case, learning may
be a central cognitive mechanism in both types of communication systems,
but the role of learning in these contexts has not been thoroughly
explored in the empirical nor theoretical literature.

Associative learning is a taxonomically widespread cognitive mechanism
that allows individuals to associate rewards with environmental stimuli
and thus behave adaptively (Heyes 2012; Macphail 1982; Staddon 2016;
Behrens et al. 2008). Theory has shown that natural selection favours
such associations in complex environments where conditions are difficult
to predict (Dridi and Lehmann 2016). Associative learning is a flexible
cognitive mechanism whose underpinnings show interspecific variation
(Enquist, Lind, and Ghirlanda 2016; Quiñones et al. 2020; Prétôt,
Bshary, and Brosnan 2016; Prétôt et al. 2021). For example, species can
have different learning rates (Hoedjes et al. 2010), some may vary these
rates during learning trials which allows them to have more flexible
learning (Leimar, Quiñones, and Bshary 2023), and some may include
future rewards in their associations (Raby et al. 2007; Osvath 2009).
Associative learning is not a single mechanism, but rather a set of
cognitive structures and processes that can vary in their scope and
complexity. Presumably, these structures and processes have been
modified by natural selection and could provide novel explanations for
behavioural variation. Nonetheless, associative learning is not often
included in the narrative of evolutionary explanations of behavioural
patterns (Fawcett, Hamblin, and Giraldeau 2013; Kamil 1983; McAuliffe
and Thornton 2015).

Computational models of evolution can overcome the lack of integration
between learning and evolution. Reinforcement learning theory
encompasses a series of computational methods inspired on the
psychological and neurological mechanisms of associative learning
(Sutton and Barto 2018). This set of algorithms allows the
implementation of biologically realistic problems, capturing the essence
of learning processes (Frankenhuis, Panchanathan, and Barto 2018;
Quiñones et al. 2020). Furthermore, these algorithms can be embedded in
evolutionary simulations to generate theoretical predictions of the
effect of learning on behavioural evolution (Leimar and McNamara 2019;
Leimar and Bshary 2022b).

Here, we present an evolutionary model where individuals use associative
learning to develop a tendency to behave either aggressively or
peacefully in the context of competition over resources, depending on a
quantitative morphological trait (\emph{i.e.} a badge indicating
quality) they perceive in their opponents. Over evolutionary time, the
size of the badge evolves as does its dependency on the individual's
quality. Under this simple set up, individuals can use the badge as a
signal of quality. We use the model to assess under what conditions one
might expect communication signals to evolve as handicaps or
conventions.

\hypertarget{the-model}{%
\subsection{The model}\label{the-model}}

We model the evolution of signals indicating individual quality in the
context of agonistic interactions. For simplicity, we consider a
population of haploid individuals with non-overlapping generations.
Individuals are born every generation with a level of quality (\(Q_i\),
where \(i\) is subscript of the population vector of size \(N\)) given
by a number between zero and one, which is drawn from a truncated normal
distribution \(N( 0.5,\sigma)\). An individual with quality \(0\) has
the lowest RHP, while an individual with quality \(1\) has the highest.
As they develop, individuals produce a phenotypic signal (badge), the
size of which depends on their quality according to a reaction norm
given by \(B_i=1/(1+e^{\alpha_i-\beta_i Q_i})\); where \(\alpha_i\) and
\(\beta_i\) are individual specific traits that determine the shape of
the reaction norm (Fig. \ref{fig:modelStruc} A). Variation in
\(\alpha_i\) and \(\beta_i\) means individuals can have either
uninformative (flat) or informative (logistic) norms. Informative
reaction norms represent a developmental program where the phenotype of
the individual is determined by the environmental conditions under which
it grows, thus causing a correlation between quality and signal. The
size of the signal is constrained to take values between zero and one.
We assume different values of these traits are given by different
alleles and are inherited from mother to offspring unless, with a small
probability (\(\mu\)), mutation changes the allelic value of the
offspring by an amount drawn from a normal distribution
\(TN(0,\sigma_\mu)\).

After birth, individuals go through a round of viability selection. The
survival probability of an individual is given by
\(s_i=1/(1+e^{-k_1-k_2(Q_i-B_i)})\), where \(k_1\) and \(k_2\) are
parameter values determining the shape of the survival function.
Importantly, if \(k_2=0\) survival probability is independent of
quality, while if \(k_2>0\) lower-quality individuals pay a higher price
for similar-sized badges, fulfilling the core assumption of the handicap
principle (Botero et al. 2010; Grafen 1990; Johnstone and Norris 1993).

Individuals who survive engage in a series of pairwise interactions
where they compete for resources. In each interaction individuals must
decide whether to escalate a fight or not. Following the classic
\emph{hawk-dove} game (Maynard-Smith 1982), if the focal individual
fights and its partner does not, then the focal gets as pay-off the
resource of value \(V\) while its partner gets nothing; if both
individuals restrain from fighting, then they split up the resource in
half. If both individuals decide to fight, then the winner takes over
the resource and the cost of fighting is split between the two players.
We further assume that the probability of wining a fight for the focal
individual depends on the difference in quality between it and its
interacting partner; specifically it is given by
\(p_{ij}=1/(1+e^{-k_3(Q_i-Q_j)})\), where \(k_3\) is a parameter
defining how strong the quality difference determines the winning
probability, and \(i\) and \(j\) denote the the position in the
population vector of the focal and its interacting partner,
respectively.

The decision of whether to escalate a fight against an interacting
partner can depend on the size of the partner's badge, with the
dependency being determined by the focal's experiences acquired through
a learning process. We implement learning using the actor-critic
approach from reinforcement learning (RL) theory (Sutton and Barto 2018;
Quiñones et al. 2020; Leimar and McNamara 2019; Leimar 2021).
Individuals estimate the reward (pay-off) expected from interacting with
partners of different badge sizes (the critic in RL terminology). After
each interaction they update the estimate of reward proportionally to
the difference between their current estimate and the observed reward
(prediction error \(\delta\)) and to the speed of learning
(\(\mathrm{A}\)). Furthermore, individuals express different
probabilities of retreating/attacking depending on the badge size of
their opponent (the actor in RL). They update the probability of
retreating/attacking depending on whether retreating/attacking leads to
an increase in the reward estimation. Thus, if a focal individual
decides to escalate a fight against an individual with a small badge and
this leads to an increase in the reward estimation, then the focal
individual will increase the probability of escalating fights with
individuals of small badges in the future. Given that badge size is a
real number between 0 and 1, there are infinitely many badge sizes.
Thus, the reward estimation, as well as the probability of
retreating/attacking, must be generalized across different values. To
implement generalization we use the linear function approximation method
based on radial basis functions (Sutton and Barto 2018). Specifically,
we pick \(c\) feature centres, which are evenly spaced values along the
badge size interval ({[}0,1{]}) where the updates are focused. For
simplicity, we keep the location of these feature centres constant and
the same for all individuals; they are stored in vector \(\symbf{b}\) .
Each of these feature centres is associated with a weight for reward
estimation and tendency to play retreat in a given interaction. The
reward estimation and the tendency to retreat are calculated (in every
interaction) as the sum of the weights associated with each feature
centre, weighted by the response triggered by the feature (Fig.
\ref{fig:modelStruc} B, dots represent the feature weights). The
response of each feature centre diminishes as a Gaussian function with
the distance between the feature centre and the badge size of the
partner (Fig. \ref{fig:modelStruc} B, grey line). Formally, the reward
estimation \(\hat{R}\) when the focal individual (\(i\)) faces
individual \(j\) is given by,\\
\begin{equation}
\hat{R}_{ij} =  \sum_{z=1}^{c} x_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

where \(x_z\) is the weight of feature centre \(z\) on the reward
estimation; and \(\theta\) is the width of the generalization function.
Similarly, the tendency to retreat is given by the sum of feature
weights associated with the actor, and the probability is obtained by
applying a logistic transformation (Fig. \ref{fig:modelStruc} B, black
line). Thus formally, the log-odds to retreat when facing individual
\(j\) is

\begin{equation}
logit(q_{ij}) =  \sum_{z=1}^{c} y_z e^{(-\frac{|B_i-b_z|}{2\theta^2})} 
\end{equation}

where \(y_z\) is the weight of feature centre \(z\) on the tendency to
retreat.

Individuals interact as the focal individual \(n\) times and the
interaction partner is chosen at random. Thus, the expected number of
interactions for each individual is \(n+\frac{n(N+1)}{N}\). After the
interaction round , individuals in the population reproduce with a
probability proportional to their total pay-off \(w_i\), which is a sum
of the baseline pay-off (\(w_0\)) and all the pay-offs obtained
throughout their life. Thus, the combination of natural selection and
genetic drift changes the distribution of values in \(\alpha\) and
\(\beta\) that segregates in the population, effectively changing the
badges expressed and the communication system.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/modelStruc-1}

}

\caption{Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size, *e.g* the size of a black band on a bird's chest. Red shows an informative reaction norm, whereas blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows that the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether this leads to an increase in the estimated reward. Panels C and D show the effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields a badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value. D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual.}\label{fig:modelStruc}
\end{figure}

\hypertarget{results}{%
\subsection{Results}\label{results}}

\hypertarget{learning-in-a-monomorphic-population}{%
\subsubsection{Learning in a monomorphic
population}\label{learning-in-a-monomorphic-population}}

We first present the outcome of simulations where we prevented the
evolution of the badge size (by setting the mutation rate to \(0\)), and
assumed that all individuals in the population display either an
uninformative or informative badge size with respect to their quality.
These simulations show what type of receiver strategy develops through a
learning process in such a monomorphic population. Figure
\ref{fig:modelStruc} (C and D) shows the receiver strategy developed
through learning; panel C is for receivers that faced uninformative
signals, and D for those facing informative ones. The simulations reveal
that when learners face uninformative signals (Fig. \ref{fig:modelStruc}
C), they modify their probability of retreating depending on their own
quality. Individuals with higher quality (red tones) have a high
probability of attacking after the learning process\\
while individuals of lower quality (blue tones) mostly retreat from
confrontations; in both cases, individuals behave equally irrespective
of the quality of their opponent. Thus, learning splits the population
of receivers into the two classic pure strategies of hawks and doves.
Given that we assumed a monomorphic population with unresponsive
reaction norms on the signalling side, the changes triggered by learning
only affected a small range of badge sizes. Specifically, all learning
occurred around a badge size of \(0.5\), because all badges in the
population are of this size (panel C). In contrast, when receivers face
informative reaction norms on the side of the signaller (Fig.
\ref{fig:modelStruc} B), receivers use information on the badge size of
their interacting partners to determine whether to retreat or attack.
The relationship is given by a threshold-like reaction norm, where the
decision to retreat or to attack depends on the quality of the receiver.
As expected, the higher the quality of the receiver, the larger the
badge size of the signaller that triggers a retreat.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/handicap-1}

}

\caption{Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\alpha$); on the right (A.2) changes in the distribution of the slope ($\beta$). Darker areas of the background correspond to values with high frequency, while orange lines in both panels show the mean of the distribution. X axis in the evolutionary dynamics are given in thousands of generations. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \ref{fig:modelStruc}. On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value. In the second half, the slope takes positive values and reaction norms evolve to produce an honest signal of quality. Correspondingly, individuals learn to react to badge size by increasing the probability of retreating with increasing badge sizes}\label{fig:handicap}
\end{figure}

\hypertarget{badges-as-handicaps}{%
\subsubsection{Badges as Handicaps}\label{badges-as-handicaps}}

When we allowed the badge size to evolve (\(\alpha\) and \(\beta\)
changed subject to natural selection and genetic drift) and the signal
worked as a handicap (\emph{i.e.} the cost of the badge was inversely
proportional to the quality of the individual), the sender code often
evolved to produce an honest signal (Fig. \ref{fig:handicap}). The
evolution of the sender code did not happen immediately after the start
of the evolutionary process. Instead, the evolutionary dynamics of the
reaction norm parameters (Fig. \ref{fig:handicap}, A) appeared to
involve a set of steps. First, the intercept of the reaction norm
(\(\alpha\)) describing the relationship between individual quality and
badge size evolved to higher values, reducing the average badge size in
the populations. This is because reaction norms segregating in the
population are flat at the beginning of the simulation; consequently,
the receiver code does not respond to the badge size and larger badges
are costlier and do not trigger lower attack probabilities. During the
first generations, the slope of the sender reaction norm (\(\beta\))
remains close to the initial value of zero. At around generation 4000,
the slope evolves toward positive values. Positive values in the slope
imply that larger badges correlate with higher quality (Fig.
\ref{fig:handicap} B). Receivers then learn to react to such
correlation, increasing the probability of retreating when facing
individuals with larger badges (Fig. \ref{fig:handicap} C). Hence,
natural selection favours larger values of \(\beta\), eventually leading
to an evolutionary equilibrium in which badge size is an honest signal
of quality mediated by the learned responses of receivers (Fig.
\ref{fig:handicap} B.4 and C.4).

The evolutionary trajectory portrayed in figure \ref{fig:handicap} A is
not the only possible outcome. If the slope of the sender reaction norms
(\(\beta\)) evolves, subject to genetic drift, toward negative values
before badges become handicaps, receivers never learn to react to the
size of badges. Thus, badges are only costly and do not provide
information about the quality of individuals (Fig. \ref{fig:noBadge}) .
Eventually, therefore, badges disappear from the population. In
contrast, when the badge does not work as a handicap but instead is
cost-free, the evolutionary process never leads to the establishment of
an honest signal. Instead, subject to genetic drift, the population
either evolves towards the disappearance of the badge or to the maximum
badge size. In either of these cases reaction norms are flat, so the
badge does not provide any information about quality.

\hypertarget{the-evolution-of-polymorphism-mediated-by-learning}{%
\subsubsection{The evolution of polymorphism mediated by
learning}\label{the-evolution-of-polymorphism-mediated-by-learning}}

The amount of information that agents are able to collect via learning
over their life time can strongly change the outcome of evolutionary
dynamics. In the simulations presented in figure \ref{fig:handicap} and
\ref{fig:noBadge}, individuals learned with high speed
(\(\mathrm{A} = 0.4\)) and interacted repeatedly along their lifetime
(2000 interactions on average). When we reduced the number of
interactions that individuals experienced over their lifetime to 300 on
average, we saw a drastic increase in the phenotypic and genetic
variation present in the population. Populations start monomorphic with
a value of zero for both the intercept and the slope of the reaction
norm, and mutations quickly build up a normal distribution around the
starting value. Within the first 2000 generations, the unimodal
distribution in the intercept (\(\alpha\)) splits into a bimodal one.
Later in evolutionary time, one of the peaks splits further, so at the
end of the evolutionary simulation the distribution of the intercept
(\(\alpha\)) in the population shows three distinct peaks. In the case
of \(\beta\), the peaks in the distribution are not so clear-cut, but
the variance of the distribution increases over time. These changes in
the distribution of the parameters of the sender reaction norm imply
that individuals can generally be classified into three distinct types
(Fig. \ref{fig:branching}). Two types express a flat reaction norm with
extreme values for the badge size, meaning that their badge size is not
informative of their quality, whereas the third type shows intermediate
badge sizes determined by individual quality (Fig. \ref{fig:branching}
B.3-4). In this simulation, the values of the slope in the informative
group are negative, which implies that badge size correlates negatively
with quality. Furthermore, the receiver reaction norms developed through
learning, particularly those of individuals of intermediate quality,
respond to the signal of their sender type by increasing the probability
of retreating from a fight with individuals with smaller badges (Fig.
\ref{fig:branching}).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/branching-1}

}

\caption{The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\alpha$ (A.1) and $\beta$ (A.2) along evolutionary time. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels. The unimodal distribution with which $\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes. These three peaks correspond to three type of reaction norms in panels B.1-3.}\label{fig:branching}
\end{figure}

A limited number of interactions has an effect on the amount of
variation in the evolving parameters when the signal follows the
handicap principle as well. In fig.~\ref{fig:honestDiv}, we show
simulations where individuals have on average 300 interactions in their
life and the cost of the signal is proportional to quality, following
the handicap principle. The evolutionary process leads to a combination
of reaction norm parameters where there is a positive correlation
between the size of the badge and the quality of the individual. This
relationship, however, is muddled by the fact that there are two
clusters of values for the intercept (\(\alpha\)) and slope (\(\beta\))
of the reaction norm in the population. Thus, there are two types of
reaction norms. One of the types has a steeper slope compared to the
other. This effect of an increased variance in the trait distribution is
not only triggered by lower number of interactions. Larger variances are
also found when we assume a lower speed of learning (data not shown).
This suggest that limits to the amount of information that individuals
acquire through learning allows the coexistence of different
communication strategies within a population, something previously shown
by Botero et al. (2010).

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/honestDiv-1}

}

\caption{The evolution of costly signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \ref{fig:handicap}. The middle panels show changes in the distribution of values for $\alpha$ and $\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels. Similarly to the evolutionary dynamics in Fig. \ref{fig:handicap}, first the badge size evolves towards its minimum value. Then, the slope evolves positive values determining an increasing reaction norm. In contrast to \ref{fig:handicap}, the normal distribution splits up into two modes. This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size.}\label{fig:honestDiv}
\end{figure}

\hypertarget{the-peaceful-the-aggressive-and-the-clever}{%
\subsubsection{The peaceful, the aggressive and the
clever}\label{the-peaceful-the-aggressive-and-the-clever}}

Our model revealed that the behaviour expressed by naive individuals
(\emph{i.e.} those who have not yet learned) imposes negative-frequency
dependent selection, allowing for the build up of genetic variation in
reaction norms. In the simulations presented so far, we assumed that
individuals start with a flat behavioural reaction norm such that they
escalate fights aggressively with a \(0.5\) probability regardless of
the badge size of the interacting partner. To assess whether such
initial conditions of the communication system had any role in the build
up of genetic variation, we ran a series of simulations varying the
initial conditions of the actor module in the learning model.
Specifically, we let naive individuals have a flat reaction norm with
either a 1) low (peaceful), 2) high (aggressive) probability of
escalating a fight or 3) a probability corresponding to the ESS of the
original hawk-dove game (which we call clever, due to information and
computation necessary to know the ESS). Fig. \ref{fig:startCond} shows
the distribution of values of the intercept (\(\alpha\)) and slope
(\(\beta\)) evolved in different replicates of the simulations. The
left-hand side panel, corresponding to peaceful naive individuals, is
the only one where the distribution of values is split in different
clusters. That is, most of the variation occurs within clusters. In
contrast, when naive individuals behave either aggressively or cleverly,
badges evolve toward minimum and maximum values, and so the variation is
driven mainly by differences among replicates. We can make sense of
these results by realizing that individuals change their naive behaviour
quickly in ranges of the badge size that are common in the population.
In contrast, an individual expressing a rare badge size will most likely
experience the naive behaviour of their partner. When the naive
behaviour is peaceful, individuals with a rare badge size have a fitness
advantage. This triggers negative-frequency dependent selection and the
evolution of different types of badge sizes. According to this
narrative, situations where individuals learn fast and interact
repeatedly will diminish the strength of frequency-dependent selection.

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/startCond-1}

}

\caption{The peaceful, the aggressive and the clever. Distribution of values of the intercept $\alpha$ and slope $\beta$ for individuals at the end of the evolutionary simulations. The panels show the three different initial conditions for the behaviour of individuals: peaceful, agressive and  "clever", see maintext for details. Colours indicate the replicate simulation. In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour. The simulations where individual disply a "clever" and agressive behaviour, before the learning process, show all the points of a single replicate cluster together, while in simulation where individuals start being peaceful al replicates show distinct groups of individuals spread throughout the x and y axis. Only in this condition there is a frequency-dependent process that favour diversity in the reaction norm.}\label{fig:startCond}
\end{figure}

The action of natural selection and genetic drift on the learning
parameters do not override the conditions for the evolutionary branching
(Dieckmann and Law 1996) described above. Given how crucial naive
behaviour is for the branching event described earlier, we tested
whether evolutionary processes could lead key parameters of the learning
process toward values that prevented the genetic variation to build up.
Figure \ref{fig:learnEvol} shows the dynamics from a set of simulations
where we allow for the coevolution of the reaction norm as well as key
parameters of the learning module, namely the speed of learning (\(A\))
and the behavioural tendency expressed by individuals prior to learning.
Top panels in figure \ref{fig:learnEvol} show that there is
diversification process in this coevolutionary scenario as well. That is
because, as is shown in the middle panels (B) of the same figure, the
learning parameters stay within the range necessary for the branching
event to take place. The speed of learning (Fig. \ref{fig:learnEvol}
B.1) stays well above zero throughout the simulation. In contrast, the
initial behavioural tendency maintain values around zero. Thus,
evolutionary processes acting on the learning parameters did not
override the conditions necessary for frequency-dependent selection to
boost genetic and phenotypic variation.

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

We presented a model that integrates learning with evolutionary
processes. In the model, the response to a signal in a communication
system is mediated by learning processes: individuals learn how to
respond to a quantitative trait of their interactive partners. We showed
that the learning process can mediate the evolution of an honest signal
under the handicap principle (Grafen 1990; Zahavi 1975). This conclusion
is not different from classical genetic models, where responses to
signals are innate to the individual. However, unlike classical genetic
models, our model shows that learning can also mediate the origin of a
signal polymorphism in the absence of the handicap principle. Under this
polymorphic equilibrium the population consist of distinct types of
individuals, with honest and dishonest signals. The amount of
information gathered through learning, as a well as the initial
conditions of the learning process, were crucial for the emergence and
maintenance of variation in the signalling system. Namely the
polymorphism required a limit in the amount of information collected by
individuals, as well as peaceful behaviour in naive individuals.
Furthermore, we showed that the conditions leading to the origin of the
polymorphism can be reached when evolutionary processes are able to
change the parameters of the learning system (the speed of learning and
the naive behaviour).

Associative learning is a powerful mechanism to learn about the world,
about social partners, and about one's abilities in a particular social
context. By associating cues and signals with fitness-relevant outcomes,
individuals collect information allowing them to improve their
reproductive potential. This is evident in classical situations such as
when animals learn to avoid food that makes them sick or use
environmental cues to find food. More recent theoretical work has
highlighted the potential role that associations can have in social
contexts such as hierarchy formation (Leimar 2021; Leimar and Bshary
2022a, 2022b) and cooperation (Leimar and McNamara 2019; Dridi and Akçay
2018). In these examples, models involve individuals who use various
sources of reward to make adaptive choices. Here we extended this logic
to the evolution of a communication system mediated by a signal of
quality. Previous evolutionary models of badges of status assumed that
individuals responded using a behavioural reaction norm, where the
opponent's badge and the individual's own quality determined behaviour
(Botero et al. 2010). However, there is no clear mechanism justifying
the assumption that individuals inherently know their own quality,
particularly relative to their peers. In our model, individuals not only
learn about the quality signal, but also learn about their own quality
relative to others. This can be seen in our simulations in the responses
developed by individuals of different quality. Because individuals learn
about their own quality, even in the absence of an honest signal, they
are able to make more adaptive choices. Specifically, lower quality
individuals refrained more from escalating fights, while higher-quality
individuals took over resources they could win in a fight. Thus, the
learning process we modelled and the response it mediates has
fitness-relevant consequences even in the absence of an honest signal.
This was confirmed by simulations where the speed of learning was
allowed to change subject to evolutionary processes. In those
simulations, natural selection always maintained learning rates above
zero. Our simulations however, only captured the effect of selection
mediated by the social game on the learning parameters. For example, we
did not account for potential costs of faster updating, or improved
performance in foraging tasks. Inter-specific variation in cognitive
abilities mediated by environmental differences has been reported
elsewhere (Sonnenberg et al. 2019). If such variation is partly mediated
by changes that affect general cognitive processes, like the speed of
learning, they could impact the outcome of communication systems like
the one we modelled.

Learning processes which involve collecting information on a
population-wide level promote frequency-dependent selection. For
example, frequency-dependence triggered by learning process was
highlighted by a classic study where live predators drive the evolution
of \emph{in silico} polymorphic prey (Bond and Kamil 2002). The key to
the evolution of prey polymorphism in such system is that predators are
better able to discriminate prey from the background when prey are
frequently encountered. Thus, prey morphs found in low frequency in a
population have a selective advantage. An analogous process emerged in
our simulations, in which the receiver individual learned the
appropriate response towards individuals with a common badge size in the
population. When an individual has a rare badge size it can be favoured
or unfavoured by the naive behaviour. When the naive behaviour is
peaceful, polymorphism is promoted, whereas when the naive behaviour is
aggressive polymorphism is prevented. These two situations show how
behaviours dependent on learning processes seem to generally trigger
frequency-dependent selection because the learning algorithm collects
more information on the more frequent values of the the trait
distribution. For simplicity, we modelled a single dimension (badge
size), yet the frequency-dependent effect of learning could potentially
be more important when individuals learn from multidimensional signals.

Originally, Rohwer (1975) proposed the idea that certain phenotypic
traits could be used by animals as status signals or signals of quality.
This idea has been tested repeatedly on the dark patches of some bird
species, such as the bibs exhibited by species in various families,
particularly sparrows. So, are bibs real signals of quality? If so, are
they handicaps or badge of status? From an evolutionary perspective it
is unclear how the stability of status signals can be maintained. One
option is that honesty is maintained by the cost of signal production;
that is, when the signal works as a handicap (Botero et al. 2010; Grafen
1990; Johnstone and Norris 1993). Alternatively, if the signal does not
carry production costs inversely proportional to quality, then it may
work as a convention. In this latter case, honesty is presumably
maintained by the aggressive reaction of receivers when the convention
is broken (Enquist, Ghirlanda, and Hurd 2010; Tibbetts and Dale 2004).
Our model captures both kinds of signals, although with some nuances.
When we assume the handicap principle, simulations result in the
evolution of honest signals mediated by the learned response. When we
assume a cost-free signal, under certain conditions the learned response
mediated the emergence of phenotypic variation in the signal. This
variation, particularly in the mid-range of the distribution,
facilitates the establishment of a convention: in this range,
individuals respond appropriately to the trait of their peer. A simple
correlation between the size os a signal and dominance rank (and
quality) is expected under the handicap principle, as it is shown here
and elsewhere (Botero et al. 2010; Grafen 1990; Johnstone and Norris
1993). However, the branching event we documented shows that conventions
can emerge in a less straightforward way when mediated by learning
processes. These nuances could potentially make sense of seemingly
contradictory evidence on the correlation between plumage traits and
quality. A case in point is that of the House Sparrow (\emph{Passer
domesticus}), a textbook example of badges of status. According to the
prevailing narrative (Senar 2006; Searcy and Nowicki 2005), the bib size
of males of this species is a signal of dominance rank and quality.
However, a recent meta-analysis called into question this narrative by
showing that the effect size of the association between bib size and
dominance rank is small and uncertain (Sánchez-Tójar et al. 2018). Our
simulations revealed that, conventions arose only in an intermediate
range of the signalling trait, yet empirical studies seeing for
correlations between phenotypic traits and quality are often performed
across the whole range of variation. Whether such a situation may apply
empirically to bib sizes in House Sparrows remains to be studied.

The focus of theoretical and empirical work on communication systems,
and particularly badges of status, is often explaining the presence and
absence of certain morphological traits within populations. Phenotypic
traits hypothesized to play a role in communication systems regularly
exhibit patterns of geographic variation within species that have long
called for explanations. One example of this is the leapfrog pattern
(Remsen 1984), whereby a certain morphological trait alternates its
presence and absence in a set of geographically adjunct populations,
likely as a consequence of selection (Cadena, Cheviron, and Funk 2011;
Márquez et al. 2020). However, it is unclear what type of ecological
process is behind these selective regimes. Given the wide variation in
outcomes found in our model, which depend on both stochasticity and
cognitive parameters which may vary in space owing to various processes,
we speculate that variation in cognition and learning could provide some
explanatory power in this respect. A first step to assess this
hypothesis would be to evaluate the way individuals in different
populations respond to novel traits (\emph{e.g.} Avendaño and Cadena
2021).

In sum, we have presented here an evolutionary model where the evolution
of a communication system is mediated through the learned responses of
receivers. This is a novel way to understand the evolution of
communication that integrates classical cognitive process in learning
with evolutionary explanations of communication. This approach
contributes to the further integration of proximate and ultimate
explanations in behavioral and evolutionary biology.

\hypertarget{supplementary-material}{%
\section{Supplementary material}\label{supplementary-material}}

\beginsupplement

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/noBadge-1}

}

\caption{The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \ref{fig:handicap}. The middle panels show changes in the distribution of values for $\alpha$ and $\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels.}\label{fig:noBadge}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=1\linewidth]{manuscript_1.0_files/figure-latex/learnEvol-1}

}

\caption{Coevolutionary dynamics of both the signal reaction norm and key parameters of the learning module. Dynamics are portrait as changes in the distribution of values assorting in the population. In the top panels (A)  evolutionary dynamics of the sender code; A.1 shows the intercept and A.2 the slope of the logistic reaction norm. Both of these parameters experience evolutionary branching, whereby in the end three distincs types are assorting in the population. The middle panels (B) show changes in the distribution of values for the speed of learning ($A$;B1) and the behavioural tendency before learning ($y_z^0$; B2) along evolutionary time. The speed of learning does not change drastically, it maintains values well above 0 (black line). In contrast, the initial behavioural tendency does not evolve away from 0 (black line). Both of these conditions favour evolutionary branching in the signal reaction norm. Bottom Panels correspond to snapshots of the sender code. Generation time of the snapshots are indicated by the grey lines in the top and middle panels.}\label{fig:learnEvol}
\end{figure}

\newpage
\small

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7857}}@{}}
\caption{Notation of the model parameters and variables}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Symbol
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Symbol
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Description
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(Q_i\) & Quality of individual \(i\) \\
\(N\) & Population size \\
\(n\) & Number of interactions as the focal individual \\
\(\sigma\) & Standard deviation of the truncated normal distribution
from which quality is drawn \\
\(B_i\) & Badge size of individual \(i\) \\
\(\alpha_i\) & Intercept of the badge size reaction norm for individual
\(i\) \\
\(\beta_i\) & Slope of the badge size reaction norm for individual
\(i\) \\
\(\mu\) & Mutation rate \\
\(\sigma_\mu\) & Standard deviation of the normal distribution from
which quality is mutations are drawn \\
\(s_i\) & Survival probability of individual \(i\) \\
\(k_1\) & Intercept of the survival function \\
\(k_2\) & Slope of the survival function \\
\(V\) & Value of the contested resource \\
\(C\) & Cost of an escalated fight \\
\(p_{ij}\) & Probability that individual \(i\) wins an escalated fight
against individual \(j\) \\
\(k_3\) & Parameter determining how important is quality defining the
probability of wining \\
\(\delta\) & Prediction error \\
\(A\) & Speed of learning \\
\(c\) & Number of feature centres \\
\(b_z\) & badge size for feature centre \(z\) \\
\(\hat{R}\) & Reward estimate \\
\(x_z\) & weight of feature centre \(z\) on the reward estimation \\
\(\theta\) & width of the generalization function \\
\(q_{ij}\) & probability that individual \(i\) retreats from a fight in
an encounter with individual \(j\) \\
\(y_z\) & weight of feature centre z on the tendency to retreat \\
\(w_i\) & total pay-off of individual \(i\) \\
\(w_0\) & base line pay-off \\
\end{longtable}

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-avendano_Territorial_2021}{}}%
Avendaño, Jorge Enrique, and Carlos Daniel Cadena. 2021. {``Territorial
Males Do Not Discriminate Between Local and Novel Plumage Phenotypes in
a Tropical Songbird Species Complex: Implications for the Role of Social
Selection in Trait Evolution.''} \emph{Behavioral Ecology and
Sociobiology} 75 (2): 37.
\url{https://doi.org/10.1007/s00265-021-02976-8}.

\leavevmode\vadjust pre{\hypertarget{ref-behrens_Associative_2008}{}}%
Behrens, Timothy E. J., Laurence T. Hunt, Mark W. Woolrich, and Matthew
F. S. Rushworth. 2008. {``Associative Learning of Social Value.''}
\emph{Nature} 456 (7219): 245--49.
\url{https://doi.org/10.1038/nature07538}.

\leavevmode\vadjust pre{\hypertarget{ref-bond_Visual_2002}{}}%
Bond, Alan B., and Alan C. Kamil. 2002. {``Visual Predators Select for
Crypticity and Polymorphism in Virtual Prey.''} \emph{Nature} 415
(6872): 609--13. \url{https://doi.org/10.1038/415609a}.

\leavevmode\vadjust pre{\hypertarget{ref-botero_Evolution_2010}{}}%
Botero, C. A, I. Pen, J. Komdeur, and F. J Weissing. 2010. {``The
Evolution of Individual Variation in Communication Strategies.''}
\emph{Evolution} 64 (11): 3123--33.

\leavevmode\vadjust pre{\hypertarget{ref-bradbury_Principles_2011}{}}%
Bradbury, Jack W., and Sandra L. Vehrencamp. 2011. \emph{Principles of
{Animal} {Communication}}. 2nd Edition. Sunderland, Mass: Sinauer
Associates is an imprint of Oxford University Press.

\leavevmode\vadjust pre{\hypertarget{ref-cadena_Testing_2011}{}}%
Cadena, C. D., Z. A. Cheviron, and W. C. Funk. 2011. {``Testing the
Molecular and Evolutionary Causes of a {`Leapfrog'} Pattern of
Geographical Variation in Coloration.''} \emph{Journal of Evolutionary
Biology} 24 (2): 402--14.
\url{https://doi.org/10.1111/j.1420-9101.2010.02175.x}.

\leavevmode\vadjust pre{\hypertarget{ref-dieckmann_Dynamical_1996}{}}%
Dieckmann, Ulf, and Richard Law. 1996. {``The Dynamical Theory of
Coevolution: A Derivation from Stochastic Ecological Processes.''}
\emph{Journal of Mathematical Biology} 34 (5): 579--612.
\url{https://doi.org/10.1007/BF02409751}.

\leavevmode\vadjust pre{\hypertarget{ref-dridi_Learning_2018}{}}%
Dridi, Slimane, and Erol Akçay. 2018. {``Learning to {Cooperate}: {The}
{Evolution} of {Social} {Rewards} in {Repeated} {Interactions}.''}
\emph{The American Naturalist} 191 (1): 58--73.
\url{https://doi.org/10.1086/694822}.

\leavevmode\vadjust pre{\hypertarget{ref-dridi_Environmental_2016}{}}%
Dridi, Slimane, and Laurent Lehmann. 2016. {``Environmental Complexity
Favors the Evolution of Learning.''} \emph{Behavioral Ecology} 27 (3):
842--50. \url{https://doi.org/10.1093/beheco/arv184}.

\leavevmode\vadjust pre{\hypertarget{ref-enquist_Signaling_2010}{}}%
Enquist, Magnus, Stefano Ghirlanda, and Peter L. Hurd. 2010.
{``Signaling.''} In \emph{Evolutionary {Behavioral} {Ecology}}. Oxford
University Press.

\leavevmode\vadjust pre{\hypertarget{ref-enquist_Power_2016}{}}%
Enquist, Magnus, Johan Lind, and Stefano Ghirlanda. 2016. {``The Power
of Associative Learning and the Ontogeny of Optimal Behaviour.''}
\emph{Royal Society Open Science} 3 (11): 160734.
\url{https://doi.org/10.1098/rsos.160734}.

\leavevmode\vadjust pre{\hypertarget{ref-fawcett_Exposing_2013}{}}%
Fawcett, Tim W., Steven Hamblin, and Luc-Alain Giraldeau. 2013.
{``Exposing the Behavioral Gambit: The Evolution of Learning and
Decision Rules.''} \emph{Behavioral Ecology} 24 (1): 2--11.
\url{https://doi.org/10.1093/beheco/ars085}.

\leavevmode\vadjust pre{\hypertarget{ref-frankenhuis_Enriching_2018}{}}%
Frankenhuis, Willem E., Karthik Panchanathan, and Andrew G. Barto. 2018.
{``Enriching Behavioral Ecology with Reinforcement Learning Methods.''}
\emph{Behavioural Processes}, February.
\url{https://doi.org/10.1016/j.beproc.2018.01.008}.

\leavevmode\vadjust pre{\hypertarget{ref-grafen_Biological_1990}{}}%
Grafen, Alan. 1990. {``Biological Signals as Handicaps.''} \emph{Journal
of Theoretical Biology} 144 (4): 517--46.
\url{https://doi.org/10.1016/S0022-5193(05)80088-8}.

\leavevmode\vadjust pre{\hypertarget{ref-guilford_Receiver_1991}{}}%
Guilford, Tim, and Marian Stamp Dawkins. 1991. {``Receiver Psychology
and the Evolution of Animal Signals.''} \emph{Animal Behaviour} 42 (1):
1--14. \url{https://doi.org/10.1016/S0003-3472(05)80600-1}.

\leavevmode\vadjust pre{\hypertarget{ref-heyes_Simple_2012}{}}%
Heyes, Cecilia. 2012. {``Simple Minds: A Qualified Defence of
Associative Learning.''} \emph{Philosophical Transactions of the Royal
Society B: Biological Sciences} 367 (1603): 2695--2703.
\url{https://doi.org/10.1098/rstb.2012.0217}.

\leavevmode\vadjust pre{\hypertarget{ref-hoedjes_Natural_2010}{}}%
Hoedjes, Katja M., H. Marjolein Kruidhof, Martinus E. Huigens, Marcel
Dicke, Louise E. M. Vet, and Hans M. Smid. 2010. {``Natural Variation in
Learning Rate and Memory Dynamics in Parasitoid Wasps: Opportunities for
Converging Ecology and Neuroscience.''} \emph{Proceedings of the Royal
Society B: Biological Sciences} 278 (1707): 889--97.
\url{https://doi.org/10.1098/rspb.2010.2199}.

\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Recognition_1997}{}}%
Johnstone, Rufus A. 1997. {``Recognition and the Evolution of
Distinctive Signatures: When Does It Pay to Reveal Identity?''}
\emph{Proceedings of the Royal Society of London. Series B: Biological
Sciences} 264 (1387): 1547--53.
\url{https://doi.org/10.1098/rspb.1997.0215}.

\leavevmode\vadjust pre{\hypertarget{ref-johnstone_Badges_1993}{}}%
Johnstone, Rufus A., and Ken Norris. 1993. {``Badges of Status and the
Cost of Aggression.''} \emph{Behavioral Ecology and Sociobiology} 32
(2): 127--34. \url{https://doi.org/10.1007/BF00164045}.

\leavevmode\vadjust pre{\hypertarget{ref-kamil_Optimal_1983}{}}%
Kamil, Alan C. 1983. {``Optimal {Foraging} {Theory} and the {Psychology}
of {Learning}.''} \emph{Integrative and Comparative Biology} 23 (2):
291--302. \url{https://doi.org/10.1093/icb/23.2.291}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Evolution_2021}{}}%
Leimar, Olof. 2021. {``The {Evolution} of {Social} {Dominance} Through
{Reinforcement} {Learning}.''} \emph{The American Naturalist} 197 (5):
560--75. \url{https://doi.org/10.1086/713758}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Reproductive_2022}{}}%
Leimar, Olof, and Redouan Bshary. 2022a. {``Reproductive Skew, Fighting
Costs and Winner--Loser Effects in Social Dominance Evolution.''}
\emph{Journal of Animal Ecology} n/a (n/a).
\url{https://doi.org/10.1111/1365-2656.13691}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Effects_2022}{}}%
---------. 2022b. {``Effects of Local Versus Global Competition on
Reproductive Skew and Sex Differences in Social Dominance Behaviour.''}
\emph{Proceedings of the Royal Society B: Biological Sciences} 289
(1987): 20222081. \url{https://doi.org/10.1098/rspb.2022.2081}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Learning_2019}{}}%
Leimar, Olof, and John M. McNamara. 2019. {``Learning Leads to Bounded
Rationality and the Evolution of Cognitive Bias in Public Goods
Games.''} \emph{Scientific Reports} 9 (1): 1--9.
\url{https://doi.org/10.1038/s41598-019-52781-7}.

\leavevmode\vadjust pre{\hypertarget{ref-leimar_Flexibility_2023}{}}%
Leimar, Olof, Andrés E. Quiñones, and Redouan Bshary. 2023.
{``Flexibility of Learning in Complex Worlds.''} bioRxiv.
\url{https://doi.org/10.1101/2023.06.12.544544}.

\leavevmode\vadjust pre{\hypertarget{ref-macphail_Brain_1982}{}}%
Macphail, Euan M. 1982. \emph{Brain and {Intelligence} in
{Vertebrates}}. Clarendon Press.

\leavevmode\vadjust pre{\hypertarget{ref-marquez_Divergence_2020}{}}%
Márquez, Roberto, Tyler P. Linderoth, Daniel Mejía-Vargas, Rasmus
Nielsen, Adolfo Amézquita, and Marcus R. Kronforst. 2020. {``Divergence,
Gene Flow, and the Origin of Leapfrog Geographic Distributions: {The}
History of Colour Pattern Variation in {Phyllobates} Poison-Dart
Frogs.''} \emph{Molecular Ecology} 29 (19): 3702--19.
\url{https://doi.org/10.1111/mec.15598}.

\leavevmode\vadjust pre{\hypertarget{ref-maynard-smith_Evolution_1982}{}}%
Maynard-Smith, John. 1982. \emph{Evolution and the {Theory} of {Games}}.
1 edition. Cambridge ; New York: Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-mcauliffe_Psychology_2015}{}}%
McAuliffe, K., and A. Thornton. 2015. {``The Psychology of Cooperation
in Animals: An Ecological Approach.''} \emph{Journal of Zoology} 295
(1): 23--35. \url{https://doi.org/10.1111/jzo.12204}.

\leavevmode\vadjust pre{\hypertarget{ref-moller_Female_1988}{}}%
Møller, Anders Pape. 1988. {``Female Choice Selects for Male Sexual Tail
Ornaments in the Monogamous Swallow.''} \emph{Nature} 332 (6165):
640--42. \url{https://doi.org/10.1038/332640a0}.

\leavevmode\vadjust pre{\hypertarget{ref-osvath_Spontaneous_2009}{}}%
Osvath, Mathias. 2009. {``Spontaneous Planning for Future Stone Throwing
by a Male Chimpanzee.''} \emph{Current Biology} 19 (5): R190--91.
\url{https://doi.org/10.1016/j.cub.2009.01.010}.

\leavevmode\vadjust pre{\hypertarget{ref-parker_Assessment_1974}{}}%
Parker, G. A. 1974. {``Assessment Strategy and the Evolution of Fighting
Behaviour.''} \emph{Journal of Theoretical Biology} 47 (1): 223--43.
\url{https://doi.org/10.1016/0022-5193(74)90111-8}.

\leavevmode\vadjust pre{\hypertarget{ref-pretot_Factors_2016}{}}%
Prétôt, Laurent, Redouan Bshary, and Sarah F. Brosnan. 2016. {``Factors
Influencing the Different Performance of Fish and Primates on a
Dichotomous Choice Task.''} \emph{Animal Behaviour} 119 (September):
189--99. \url{https://doi.org/10.1016/j.anbehav.2016.06.023}.

\leavevmode\vadjust pre{\hypertarget{ref-pretot_Comparative_2021}{}}%
Prétôt, Laurent, Jennifer Mickelberg, Jodi Carrigan, Tara Stoinski,
Redouan Bshary, and Sarah F. Brosnan. 2021. {``Comparative Performance
of Orangutans ({Pongo} Spp.), Gorillas ({Gorilla} Gorilla Gorilla), and
Drills ({Mandrillus} Leucophaeus), in an Ephemeral Foraging Task.''}
\emph{American Journal of Primatology} 83 (1): e23212.
\url{https://doi.org/10.1002/ajp.23212}.

\leavevmode\vadjust pre{\hypertarget{ref-quinones_Negotiation_2016}{}}%
Quiñones, Andrés E., G. Sander van Doorn, Ido Pen, Franz J. Weissing,
and Michael Taborsky. 2016. {``Negotiation and Appeasement Can Be More
Effective Drivers of Sociality Than Kin Selection.''} \emph{Phil. Trans.
R. Soc. B} 371 (1687): 20150089.
\url{https://doi.org/10.1098/rstb.2015.0089}.

\leavevmode\vadjust pre{\hypertarget{ref-quinones_Reinforcement_2020}{}}%
Quiñones, Andrés E., Olof Leimar, Arnon Lotem, and Redouan Bshary. 2020.
{``Reinforcement {Learning} {Theory} {Reveals} the {Cognitive}
{Requirements} for {Solving} the {Cleaner} {Fish} {Market} {Task}.''}
\emph{The American Naturalist} 195 (4): 664--77.
\url{https://doi.org/10.1086/707519}.

\leavevmode\vadjust pre{\hypertarget{ref-raby_Planning_2007}{}}%
Raby, C. R., D. M. Alexis, A. Dickinson, and N. S. Clayton. 2007.
{``Planning for the Future by Western Scrub-Jays.''} \emph{Nature} 445
(7130): 919. \url{https://doi.org/10.1038/nature05575}.

\leavevmode\vadjust pre{\hypertarget{ref-remsen_High_1984}{}}%
Remsen, J. V. 1984. {``High {Incidence} of "{Leapfrog}" {Pattern} of
{Geographic} {Variation} in {Andean} {Birds}: {Implications} for the
{Speciation} {Process}.''} \emph{Science} 224 (4645): 171--73.
\url{https://doi.org/10.1126/science.224.4645.171}.

\leavevmode\vadjust pre{\hypertarget{ref-rohwer_Social_1975a}{}}%
Rohwer, Sievert. 1975. {``The {Social} {Significance} of {Avian}
{Winter} {Plumage} {Variability}.''} \emph{Evolution} 29 (4): 593--610.
\url{https://doi.org/10.2307/2407071}.

\leavevmode\vadjust pre{\hypertarget{ref-sanchez-tojar_Metaanalysis_2018a}{}}%
Sánchez-Tójar, Alfredo, Shinichi Nakagawa, Moisès Sánchez-Fortún,
Dominic A Martin, Sukanya Ramani, Antje Girndt, Veronika Bókony, et al.
2018. {``Meta-Analysis Challenges a Textbook Example of Status
Signalling and Demonstrates Publication Bias.''} Edited by Diethard
Tautz and Tim Parker. \emph{eLife} 7 (November): e37385.
\url{https://doi.org/10.7554/eLife.37385}.

\leavevmode\vadjust pre{\hypertarget{ref-searcy_Evolution_2005}{}}%
Searcy, William A., and Stephen Nowicki. 2005. \emph{The {Evolution} of
{Animal} {Communication}: {Reliability} and {Deception} in {Signaling}
{Systems}}. Princeton: Princeton University Press.

\leavevmode\vadjust pre{\hypertarget{ref-3062aab1-6f44-361c-a548-6bc040ba8ca2}{}}%
Senar, Juan Carlos. 2006. {``Color Displays as Intrasexual Signals of
Aggression and Dominance.''} In \emph{Bird Coloration, Volume 2:
{Function} and Evolution}, 87--136. Harvard University Press.
\url{http://www.jstor.org/stable/j.ctv22jnr8k.6}.

\leavevmode\vadjust pre{\hypertarget{ref-sonnenberg_Natural_2019}{}}%
Sonnenberg, Benjamin R., Carrie L. Branch, Angela M. Pitera, Eli Bridge,
and Vladimir V. Pravosudov. 2019. {``Natural {Selection} and {Spatial}
{Cognition} in {Wild} {Food}-{Caching} {Mountain} {Chickadees}.''}
\emph{Current Biology} 29 (4): 670--676.e3.
\url{https://doi.org/10.1016/j.cub.2019.01.006}.

\leavevmode\vadjust pre{\hypertarget{ref-staddon_Adaptive_2016}{}}%
Staddon, J. E. R. 2016. \emph{Adaptive {Behavior} and {Learning}}. 2
edition. Cambridge: Cambridge University Press.

\leavevmode\vadjust pre{\hypertarget{ref-sutton_Reinforcement_2018}{}}%
Sutton, Richard S., and Andrew G. Barto. 2018. \emph{Reinforcement
{Learning}: {An} {Introduction}}. Edited by Francis Bach. Second edition
edition. Cambridge, MA: A Bradford Book.

\leavevmode\vadjust pre{\hypertarget{ref-tibbetts_Socially_2004}{}}%
Tibbetts, Elizabeth A., and James Dale. 2004. {``A Socially Enforced
Signal of Quality in a Paper Wasp.''} \emph{Nature} 432 (7014): 218--22.
\url{https://doi.org/10.1038/nature02949}.

\leavevmode\vadjust pre{\hypertarget{ref-zahavi_Mate_1975}{}}%
Zahavi, Amotz. 1975. {``Mate Selection----{A} Selection for a
Handicap.''} \emph{Journal of Theoretical Biology} 53 (1): 205--14.
\url{https://doi.org/10.1016/0022-5193(75)90111-3}.

\end{CSLReferences}

\end{document}